---
title: Exploring the Factors of Song Popularity
author: by Louhavidandros
output: github_document
---

```{r load-packages, include = FALSE}
library(tidyverse)
library(dplyr)
library(ggridges)
library(fmsb)
library(patchwork)
library(knitr)
```

```{r load-data, include = FALSE}
spotify_songs <- read_csv(here::here("data/Spotify.csv"))
```

```{r clean-data, include = FALSE}
clean_songs <- spotify_songs %>%
  filter(duplicated(track_name) == FALSE)
clean_full <- spotify_songs %>%
  group_by(playlist_genre) %>%
  filter(duplicated(track_name) == FALSE)
```

```{r popularity_level_clean_data, include = FALSE}
clean_edm <- clean_full %>%
  filter(playlist_genre == "edm") %>%
  mutate(popularity_level = case_when(
    track_popularity <= 17 ~ "Very Low",
    track_popularity > 17 & track_popularity <= 35 ~ "Low",
    track_popularity > 35 & track_popularity < 49 ~ "Fairly High",
    track_popularity >= 49 & track_popularity <= 100 ~ "High"
  ))

clean_latin <- clean_full %>%
  filter(playlist_genre == "latin") %>%
  mutate(popularity_level = case_when(
    track_popularity <= 31 ~ "Very Low",
    track_popularity > 31 & track_popularity <= 48 ~ "Low",
    track_popularity > 48 & track_popularity < 63 ~ "Fairly High",
    track_popularity >= 63 & track_popularity <= 100 ~ "High"
  ))

clean_rnb <- clean_full %>%
  filter(playlist_genre == "r&b") %>%
  mutate(popularity_level = case_when(
    track_popularity <= 21 ~ "Very Low",
    track_popularity > 21 & track_popularity <= 43 ~ "Low",
    track_popularity > 43 & track_popularity < 61 ~ "Fairly High",
    track_popularity >= 61 & track_popularity <= 100 ~ "High"
  ))

clean_rap <- clean_full %>%
  filter(playlist_genre == "rap") %>%
  mutate(popularity_level = case_when(
    track_popularity <= 30 ~ "Very Low",
    track_popularity > 30 & track_popularity <= 47 ~ "Low",
    track_popularity > 47 & track_popularity < 59 ~ "Fairly High",
    track_popularity >= 59 & track_popularity <= 100 ~ "High"
  ))

clean_pop <- clean_full %>%
  filter(playlist_genre == "pop") %>%
  mutate(popularity_level = case_when(
    track_popularity <= 30 ~ "Very Low",
    track_popularity > 30 & track_popularity <= 50 ~ "Low",
    track_popularity > 50 & track_popularity < 65 ~ "Fairly High",
    track_popularity >= 65 & track_popularity <= 100 ~ "High"
  )) 

clean_rock <- clean_full %>%
  filter(playlist_genre == "rock") %>%
  mutate(popularity_level = case_when(
    track_popularity <= 24 ~ "Very Low",
    track_popularity > 24 & track_popularity <= 45 ~ "Low",
    track_popularity > 45 & track_popularity < 61 ~ "Fairly High",
    track_popularity >= 61 & track_popularity <= 100 ~ "High"
  )) 

all_clean <- clean_rock %>%
  rbind(clean_rnb, clean_rap, clean_pop, clean_latin, clean_edm)
```
# Summary

We chose to use the `spotify_songs`dataset. variables, which range from information about the song such as name, as well as numerical variables which look at aspects of the songs character such as its energy. 
Our goal was to explore every genre of music and see how the numerical variables affected a song's popularity. We then wanted to compare this across genres to see whether they had different contributing factors .
Finally we looked at how the characteristics of popular songs have changed over time, and predicted what characteristics a popular song released next year might have.

Firstly, the data set was cleaned by filtering for each genre and removing all repeated songs.
We did a genre-by-genre analysis of each variable by creating scatter plots with a `geom_smooth` line, as well as a linear model which would use every numerical variable as an explanatory variable to try and predict track popularity. 
The scatter plots were unsuccessful, and the linear model was not reliable enough for consideration.

## Density and Ridge Plots

Consequently, we created a new variable - `popularity_level` - by categorising song's popularity based on which quartile they fell into. This was done separately for each genre, this way there was a roughly even number of songs in each popularity level in each genre.
For every genre we then looked at only the "High" and "Very Low" brackets of popularity, and for each variable we created a density plot.


```{r pop-density-example, echo = FALSE}
clean_pop %>%
  filter(popularity_level %in% c("High", "Very Low")) %>%
  ggplot(aes(x = danceability, color = popularity_level)) +
  geom_density(adjust = 2) +
    labs(x = "Danceability",
       y = "Frequency",
       color = "Popularity Level") +
  theme_minimal()
```

This shows the distribution of danceability levels of pop songs in the highest and lowest popularity brackets. The range of values where the red line is higher than the blue was defined as the optimal range for song popularity, as this is where if you picked a pop song at random you would have a higher chance of it being very popular than being very unpopular. 

Ridge plots were made for every variable so that comparisons could be made:

```{r ridge-plot, echo = FALSE}
all_clean %>%
  filter(popularity_level %in% c("High", "Very Low")) %>%
  ggplot(aes(x = danceability,
             y = playlist_genre,
             fill = popularity_level,
             color = popularity_level)) +
   geom_density_ridges(alpha = 0.2, bandwidth = 0.045) +
  labs(x = "Danceability",
       y = NULL,
       fill = "Popularity Level",
       title = "Distibution of Danceability Across Genres",
       subtitle = "Grouped by Most and Least Popular Songs"
       ) +
  guides(color = FALSE) +
  scale_fill_manual(values = c(
    "Very Low" = "#008dfd",
    "High" = "#fd0000"
  )) +
  theme_minimal()
```

Rock is on average the least danceable as its distribution is furthest left.
Danceability has a negligible effect on the likelihood of popularity for EDM music, and only a small effect for R&B. For pop, rap and latin  and rock, a high danceability leads to a higher likelihood of a song being popular.

These graphs were made for every variable which showed useful information.

## Multiple Variable Analysis

After having looked at individual variables, we explored a combination of variables - danceability, energy and valence:

```{r pop-multiple-variable-analysis-1, echo = FALSE}
mvg_pop <- clean_pop %>%
  mutate(dance_energy_val = danceability*energy*valence) %>%
  ggplot(aes(x = popularity_level, y = dance_energy_val, colour = popularity_level)) +
  geom_col() +
  labs(
    title = "Pop",
    y = "Count",
    x = "Popularity Level") +
  theme(legend.position = "none")

mvg_edm <- clean_edm %>%
  mutate(dance_energy_val = danceability*energy*valence) %>%
  ggplot(aes(x = popularity_level, y = dance_energy_val, colour = popularity_level)) +
  geom_col() +
  labs(
    title = "EDM",
    y = "Count",
    x = "Popularity Level") +
  theme(legend.position = "none")

mvg_rap <- clean_rap %>%
  mutate(dance_energy_val = danceability*energy*valence) %>%
  ggplot(aes(x = popularity_level, y = dance_energy_val, colour = popularity_level)) +
  geom_col() +
  labs(
    title = "Rap",
    y = "Count",
    x = "Popularity Level") +
  theme(legend.position = "none")

mvg_rock <- clean_rock %>%
  mutate(dance_energy_val = danceability*energy*valence) %>%
  ggplot(aes(x = popularity_level,y = dance_energy_val, colour = popularity_level)) +
  geom_col() +
  labs(
    title = "Rock",
    y = "Count",
    x = "Popularity Level") +
  theme(legend.position = "none")
  

 (mvg_edm | mvg_rap)/(mvg_pop | mvg_rock)
```


```{r multiple-variable-analysis-2, echo = FALSE}
mvg_latin <- clean_latin  %>%
  mutate(dance_energy_val = danceability*energy*valence) %>%
  ggplot(aes(x = popularity_level, y = dance_energy_val, colour = popularity_level)) +
  geom_col() +
  labs(
    title = "Latin",
    y = "Count",
    x = "Popularity Level") +
  theme(legend.position = "none")

mvg_rnb <- clean_rnb %>%
  mutate(dance_energy_val = danceability*energy*valence)  %>%
  ggplot(aes(x = popularity_level, y = dance_energy_val, colour = popularity_level)) +
  geom_col() +
  labs(
    title = "R&B",
    y = "Count",
    x = "Popularity Level") +
  theme(legend.position = "none")

mvg_rnb+mvg_latin
```

 As shown in graphs, the songs in the "High" bracket for popularity also had the highest value for this new variable in 4 out of the 6 genres. The 2 genres in which "High" wasn't were R&B and Latin; for Latin however it was a very close second. This shows that popular songs in general, throughout most genres, are more danceable, energetic and happy.

## Changes in Trends Over Time

```{r Timeline, echo=FALSE, message=FALSE}
all_clean %>%
  filter(popularity_level == "High") %>%
  mutate(track_album_release_year = as.integer(str_sub(track_album_release_date, end = 4))) %>%
  filter(track_album_release_year >= 1970) %>%
  group_by(track_album_release_year) %>%
  summarise(acousticness = mean(acousticness), danceability = mean(danceability), energy = mean(energy), liveness = mean(liveness), 
            speechiness = mean(speechiness), valence = mean(valence)) %>%
  pivot_longer(cols = c(danceability, energy, speechiness, acousticness,
         liveness, valence)) %>%
  ggplot(aes(x = track_album_release_year, y = value, colour = name)) +
  geom_point()+
  geom_smooth() +
  theme_minimal() +
  facet_wrap(~name) +
  labs(x = "Release year",
       y = "Value",
       colour = "Variables",
       title = "Mean Value of Variables over Time",
       subtitle = "1970 - 2020") +
theme(axis.text.x = element_text(angle=45, hjust = 1))
```

Our exploration of changes in trends over time led us to create models which predict the mean of each numeric variable using the release year as the explanatory. When we plotted the data, we saw patterns emerge which we could then use to predict the characteristics of next years songs.

## Conclusion

Whilst we did not achieve our goal by the expected means, we successfully identified many characteristics which would give a song the highest likelihood of it being popular, which are outlined in our presentation. 

```{r, echo = FALSE}
model_table <- tibble(acousticness = 0.0874, danceability = 0.6767, 
                      energy = 0.7331, instrumentalness = 0.0097, liveness = 0.1967, valence = 0.8828)
kable(head(model_table), format = "html")
```

## Evaluation
We did our best to make the data set as appropriate as possible, however there were a few problems with our data set which we could not change and that may have lead to inaccuracies in our result:

The dataset contained roughly 28,000 songs, a small fraction of the songs released in the past 50 years.

Old songs released long before Spotify was released have their popularity judged by modern standards rather than the standards of their time.

The data categorized a song's genre based on the genre of the playlist it was in which lead to some song's genres being mislabeled. 



## Presentation

Our presentation can be found [here](presentation/presentation.html).

## Data 

Thompson, C, Parry, J, Phipps, D, Wolff, T 2020, Spotify Songs, electronic dataset, R for Data Science Online Learning Community/ TidyTuesday, viewed 2 November 2020, <https://github.com/rfordatascience/tidytuesday/blob/master/data/2020/2020-01-21/readme.md>

## References

Thompson, C, Parry, J, Phipps, D, Wolff, T 2020, Spotify Songs, electronic dataset, R for Data Science Online Learning Community/ TidyTuesday, viewed 2 November 2020, <https://github.com/rfordatascience/tidytuesday/blob/master/data/2020/2020-01-21/readme.md>
