---
title: Exploring the Factors of Song Popularity
author: by Louhavidandros
output: github_document
---

```{r load-packages, include = FALSE}
library(tidyverse)
library(dplyr)
library(ggridges)
library(fmsb)
library(ggradar)
```

```{r load-data, include = FALSE}
spotify_songs <- read_csv(here::here("data/Spotify.csv"))
```

```{r clean-data, include = FALSE}
clean_songs <- spotify_songs %>%
  filter(duplicated(track_name) == FALSE)
clean_full <- spotify_songs %>%
  group_by(playlist_genre) %>%
  filter(duplicated(track_name) == FALSE)
```

```{r popularity_level_clean_data, include = FALSE}
clean_edm <- clean_full %>%
  filter(playlist_genre == "edm") %>%
  mutate(popularity_level = case_when(
    track_popularity <= 17 ~ "Very Low",
    track_popularity > 17 & track_popularity <= 35 ~ "Low",
    track_popularity > 35 & track_popularity < 49 ~ "Fairly High",
    track_popularity >= 49 & track_popularity <= 100 ~ "High"
  ))

clean_latin <- clean_full %>%
  filter(playlist_genre == "latin") %>%
  mutate(popularity_level = case_when(
    track_popularity <= 31 ~ "Very Low",
    track_popularity > 31 & track_popularity <= 48 ~ "Low",
    track_popularity > 48 & track_popularity < 63 ~ "Fairly High",
    track_popularity >= 63 & track_popularity <= 100 ~ "High"
  ))

clean_rnb <- clean_full %>%
  filter(playlist_genre == "r&b") %>%
  mutate(popularity_level = case_when(
    track_popularity <= 21 ~ "Very Low",
    track_popularity > 21 & track_popularity <= 43 ~ "Low",
    track_popularity > 43 & track_popularity < 61 ~ "Fairly High",
    track_popularity >= 61 & track_popularity <= 100 ~ "High"
  ))

clean_rap <- clean_full %>%
  filter(playlist_genre == "rap") %>%
  mutate(popularity_level = case_when(
    track_popularity <= 30 ~ "Very Low",
    track_popularity > 30 & track_popularity <= 47 ~ "Low",
    track_popularity > 47 & track_popularity < 59 ~ "Fairly High",
    track_popularity >= 59 & track_popularity <= 100 ~ "High"
  ))

clean_pop <- clean_full %>%
  filter(playlist_genre == "pop") %>%
  mutate(popularity_level = case_when(
    track_popularity <= 30 ~ "Very Low",
    track_popularity > 30 & track_popularity <= 50 ~ "Low",
    track_popularity > 50 & track_popularity < 65 ~ "Fairly High",
    track_popularity >= 65 & track_popularity <= 100 ~ "High"
  )) 

clean_rock <- clean_full %>%
  filter(playlist_genre == "rock") %>%
  mutate(popularity_level = case_when(
    track_popularity <= 24 ~ "Very Low",
    track_popularity > 24 & track_popularity <= 45 ~ "Low",
    track_popularity > 45 & track_popularity < 61 ~ "Fairly High",
    track_popularity >= 61 & track_popularity <= 100 ~ "High"
  )) 

all_clean <- clean_rock %>%
  rbind(clean_rnb, clean_rap, clean_pop, clean_latin, clean_edm)
```
# Summary

We chose to use the `spotify_songs` which gets its data from the `spotifyr`package, whose data comes from Spotify. The data set contains `r nrow(spotify_songs)` observations each of which is a song, and `r ncol(spotify_songs)` variables, which range from information about the song such as name, artist and genre as well as a number of numerical variables which each look at aspects of the songs character such as its energy, its danceability or its popularity. 
Our goal was to explore every genre of music in the data set and see for each genre how the numerical variables affected a song's popularity; once we had this we then wanted to compare this across the genres to see whether they had different factors that contributed to making songs popular.
Finally we planned to look at how the characteristics of popular songs have changed over time, and to predict what characteristics a popular song might have if it was released next year.

Prior to conducting any analysis the data set was cleaned as best we could by filtering for each genre and removing all repeated songs (there were just shy of 5000 songs which appeared in more than one playlist in a given genre). This way songs could appear multiple times if their genre is ambiguous, but repeats will not impact the analysis of one genre by itself.

To begin with, we did some very general analysis and looked at which genres and subgenres of music were the most popular overall, we then decided to do a genre by genre analysis of each variable by creating scatter plots with a `geom_smooth` line, as well as a linear model which would use every numerical variable as an explanatory variable to try and predict track popularity. 
The scatter plots were unsuccessful however, and the linear model we produced had an R squared value of 0.09 and an rmse value of 22.5 which means that either the model does not explain the variable and another type of model should have been used, or that the variables may not explain the data well. However, since we have not learned other models we decided to examine in further detail the second case.

## Density and Ridge Plots

Consequently, we changed the way we were approaching the problem. To begin with, we created a new variable - `popularity_level` - by categorising song's popularity as high, fairly high, low and very low based on which quartile they fell into. This was done separately for each genre, this way there was a roughly even number of songs in each popularity level in each genre.
For every genre we then looked at only the "High" and "Very Low" brackets of popularity, and for each variable we created a density plot that showed what values were most frequent in each popularity level. For example:


```{r pop-density-example, echo = FALSE}
clean_pop %>%
  filter(popularity_level %in% c("High", "Very Low")) %>%
  ggplot(aes(x = danceability, color = popularity_level)) +
  geom_density(adjust = 2) +
    labs(x = "Danceability",
       y = "Frequency",
       color = "Popularity Level") +
  theme_minimal()
```

This graph shows the distribution of danceability levels of pop songs in the highest and lowest popularity brackets. The range of values where the red line is higher than the blue was defined as the optimal range for song popularity, as this is where if you picked a pop song at random you would have a higher chance of it being very popular than being very unpopular. 

These density graphs were made for every variable for each genre and once these were created and we found which variables had interesting information, we used ridge plots to be able to compare the distributions across genres. For example:

```{r ridge-plot, echo = FALSE}
all_clean %>%
  filter(popularity_level %in% c("High", "Very Low")) %>%
  ggplot(aes(x = danceability,
             y = playlist_genre,
             fill = popularity_level,
             color = popularity_level)) +
   geom_density_ridges(alpha = 0.2, bandwidth = 0.045) +
  labs(x = "Danceability",
       y = NULL,
       fill = "Popularity Level",
       title = "Distibution of Danceability Across Genres",
       subtitle = "Grouped by Most and Least Popular Songs"
       ) +
  guides(color = FALSE) +
  scale_fill_manual(values = c(
    "Very Low" = "#008dfd",
    "High" = "#fd0000"
  )) +
  theme_minimal()
```

Not only do we gather that on average rock is the least danceable, but we can also see the variation in the optimal range for each genre; in all genres except edm, there is a threshold above which you are in the optimal range for danceability. This threshold is highest in R&B, however there is another optimal range which is at a much lower value but has a greater likelihood of a song being popular. In rap, pop, latin and rock however a higher danceability will result in a higher likelihood of a song being popular. In EDM, danceability has a negligible effect on the likelihood of a song being popular. These graphs were made for every variable which showed useful information.

## Multiple Variable Analysis
After having looked at individual variables, we explored a combination of variables. The combination of danceability, valence and energy into one variable shows some interesting results when genres are split into popularity levels. In this case the 3 variables were multiplied together to return a variable that measures how happy, danceable and energetic a song is. As shown in graphs, the songs in the "High" bracket for popularity also had the highest value for this new variable in 4 out of the 6 genres. The 2 genres in which "High" wasn't were R&B and Latin; for Latin however it was a very close second and while the "High" group came last in R&B, this can be explained by the nature of R&B as a genre. In summary while this doesn’t explicitly show that the more danceable, energetic and happy the song is the more likely it is to be popular, it does show that popular songs in general, throughout most genres, are more danceable, energetic and happy.

## Changes in Trends Over Time

Finally we decided to explore how the trends of popular songs have changed over time. Models were fitted to predict each numeric variable using the song release date as the explanatory variable to see whether there was any relationship between when a song was released and the level of a certain variable it had.

## Conclusion

Whilst we did not achieve our goal by the expected means, we successfully identified many characteristics which would give a song the highest likelihood of it being popular. For example, a song with a duration between 180-220 seconds has a much higher chance of being popular in all genres except rock. 




Write-up of your project and findings go here. 
Think of this as the text of your presentation. 
The length should be roughly 5 minutes when read out loud.
Although pacing varies, a 5-minute speech is roughly 750 words. 
To use the word count addin, select the text you want to count the words of (probably this is the 
Summary section of this document, go to Addins, and select the `Word count` addin).
This addin counts words using two different algorithms, but the results should be similar and as long as you're in the ballpark of 750 words, you're good! 
The addin will ignore code chunks and only count the words in prose.

You can also load your data here and present any analysis results / plots, but I strongly urge you to keep that to a minimum (maybe only the most important graphic, if you have one you can choose). 
And make sure to hide your code with  `echo = FALSE` unless the point you are trying to make is about the code itself.
Your results with proper output and graphics go in your presentation, this space is for a brief summary of your project.


## Presentation

Our presentation can be found [here](presentation/presentation.html).

## Data 

Include a citation for your data here. 
See http://libraryguides.vu.edu.au/c.php?g=386501&p=4347840 for guidance on proper citation for datasets. 
If you got your data off the web, make sure to note the retrieval date.

## References

List any references here. You should, at a minimum, list your data source.
