---
title: "Project proposal"
author: "Louhavidandros"
output: github_document
editor_options: 
  chunk_output_type: console
---

```{r load-packages, message = FALSE}
library(tidyverse)
library(broom)
```

## 1. Introduction

We have decided that for our project, it would be interesting to try and find what makes a song popular in its genre. For this, we will use spotify_songs which gets its data from the spotifyr package which in turn gets its data from spotify.

## 2. Data
```{r}
spotify_songs <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-01-21/spotify_songs.csv')
```



## 3. Data analysis plan

Since the aim of the project is to examine what aspects make a song popular the response variable will always be the popularity of a song (track_popularity). Each genre of music will be studied separately, using all the numeric variables as well as the artist of each song. The comparison groups are the genres of music.

```{r}
spotify_songs %>%
  count(playlist_genre)
```

```{r}
spotify_songs %>%
  group_by(playlist_genre) %>%
  summarise(min_pop = min(track_popularity), mean_pop = mean(track_popularity), max_pop = max(track_popularity)) %>%
  ggplot(aes(x = mean_pop, y = playlist_genre, fill = playlist_genre)) + 
  geom_col() + theme_minimal() + 
  labs(title = "Average Popularity of Each Genre", x = "Mean Popularity", y = NULL) +
  guides(fill = FALSE)
```

From the plot we can see that on average, pop music is the most popular and Latin music is close behind (possibly from large quantities of listeners from Spanish-speaking countries). Electronic Dance Music is the least popular on average.

