---
title: "Alexandros - Rap - Latin - R&B "
output: github_document
editor_options: 
  chunk_output_type: console
---
```{r load-packages}
library(tidyverse)
library(here)
library(lubridate)
```

```{r read-data}
spotify_songs <- read_csv(here('data/Spotify.csv'))
```

```{r add-levels}
spotify_songs <- spotify_songs %>%
mutate(popularity_level = case_when(
track_popularity <= 31 ~ "Very Low",
track_popularity > 31 & track_popularity <= 52 ~ "Low",
track_popularity > 52 & track_popularity <= 68 ~ "Fairly High",
track_popularity > 68 & track_popularity <= 100 ~ "High"
)) 
```

##RAP

```{r summarry-rap}
spotify_songs %>%
filter(playlist_genre == "rap") %>%
summarise(min_pop = min(track_popularity), iqr_pop = IQR(track_popularity), med_pop = median(track_popularity), max_pop = max(track_popularity), quantile(track_popularity))
```

##Check popularity of a song with key and mode

```{r popularity-vs-key-for-major-rap}
spotify_songs %>%
  filter(playlist_genre == "rap") %>%
  filter(mode == 1) %>%
  group_by(key) %>%
  ggplot(aes(y = track_popularity, x = key)) +
  geom_col() +
  facet_wrap(.~ popularity_level)
```

```{r popularity-vs-key-for-minor-rap}
spotify_songs %>%
  filter(playlist_genre == "rap") %>%
  filter(mode == 0) %>%
  group_by(key) %>%
  ggplot(aes(y = track_popularity, x = key)) +
  geom_col() +
  facet_wrap(.~ popularity_level)
```

```{r popularity-vs-key-for-minor-rap-j}
spotify_songs %>%
  filter(playlist_genre == "rap") %>%
  filter(mode == 0) %>%
  group_by(key) %>%
  ggplot(aes(x = key)) +
  geom_histogram(binwidth = 0.5) +
  facet_wrap(.~ popularity_level)
```


c = 0, c# = 1, d = 2, d# = 3,E = 4, F = 5, F# = 6, G = 7, G# = 8, A = 9, A# = 10, B = 11, NA = -1 

##Check popularity of a song with instrumentalness

```{r popularity-vs-instrumentalness-rap}
spotify_songs %>%
  filter(playlist_genre == "rap") %>%
  ggplot(aes(y = track_popularity, x = instrumentalness)) +
  geom_hex() +
  facet_wrap(.~ popularity_level)
```


##Check popularity of a song with duration

```{r popularity-vs-duration-rap}
spotify_songs %>%
  filter(playlist_genre == "rap") %>%
  mutate(duration_minutes = duration_ms/(1000*60)) %>%
  ggplot(aes(x = duration_minutes, y = track_popularity)) +
  geom_point() +
  geom_smooth() 
  
```

```{r most-popular-duration-rap}
spotify_songs %>%
  filter(playlist_genre == "rap",track_popularity > 68) %>%
  summarise(average_duration_rap = mean(duration_ms)/(1000*60))
  
```
 We can see that between 2.5-3.3 there are the most popular songs, which makes sense since very long songs tend not to be so popular.
 
 
##Check popularity of a song with tempo
```{r popularity-vs-tempo-rap}
spotify_songs %>%
  filter(playlist_genre == "rap") %>%
  ggplot(aes(x = tempo, y = track_popularity)) +
  geom_point() +
  geom_smooth() 
```

```{r most-popular-tempo-rap} 
#CHECK VARIATION
spotify_songs %>%
  filter(playlist_genre == "rap",track_popularity > 68) %>%
  summarise(average_tempo = mean(tempo))
  
```


##Check popularity of a song with speechiness

```{r popularity-vs-speechiness-rap}
spotify_songs %>%
  filter(playlist_genre == "rap") %>%
  ggplot(aes(y = track_popularity, x = speechiness)) +
  geom_col() 
```

```{r popularity-vs-speechiness-by-popularity-lvl-rap}
spotify_songs %>%
  filter(playlist_genre == "rap") %>%
  ggplot(aes(y = track_popularity, x = speechiness)) +
  geom_col() +
  facet_wrap(.~ popularity_level)
```

```{r most-popular-speechnes-lvl-rap}
spotify_songs %>%
  filter(playlist_genre == "rap",track_popularity > 68) %>%
  summarise(average_speechness = mean(speechiness))
```


Most popular songs are below 0.25. "Speechiness detects the presence of spoken words in a track. The more exclusively speech-like the recording (e.g. talk show, audio book, poetry), the closer to 1.0 the attribute value. Values above 0.66 describe tracks that are probably made entirely of spoken words. Values between 0.33 and 0.66 describe tracks that may contain both music and speech, either in sections or layered, including such cases as rap music. Values below 0.33 most likely represent music and other non-speech-like tracks."

##Check popularity of a song with mode

```{r popularity-vs-mode-rap}
spotify_songs %>%
  filter(playlist_genre == "rap") %>%
  ggplot(aes(y = track_popularity,  x = mode)) +
  geom_col()
```

```{r popularity-vs-mode-by-popularity-lvl-rap}
spotify_songs %>%
  filter(playlist_genre == "rap") %>%
  ggplot(aes(y = track_popularity,  x = mode)) +
  geom_col() +
  facet_wrap(.~ popularity_level)
```

##Check popularity of a song with danceability

```{r popularity-vs-danceability-rap}
spotify_songs %>%
  filter(playlist_genre == "rap") %>%
  ggplot(aes(y = track_popularity,  x = danceability)) +
  geom_col() 
```
 
```{r popularity-vs-danceability-by-popularity-lvl-rap}
spotify_songs %>%
  filter(playlist_genre == "rap") %>%
  ggplot(aes(y = track_popularity,  x = danceability)) +
  geom_col() +
  facet_wrap(.~ popularity_level)
```

```{r most-popular-danceability-rap}
spotify_songs %>%
  filter(playlist_genre == "rap",track_popularity > 68) %>%
  summarise(average_danceability = mean(danceability))
```


##Check popularity of a song with release date

```{r popularity-vs-releasedate-rap}
spotify_songs %>%
  filter(playlist_genre == "rap") %>%
  mutate(release_date = as_date(track_album_release_date)) %>%
  ggplot(aes(y = track_popularity,  x = release_date)) +
  geom_col()
```

##Check popularity of a song with energy

```{r popularity-vs-energy-rap}
spotify_songs %>%
  filter(playlist_genre == "rap") %>%
  ggplot(aes(y = track_popularity,  x = energy)) +
  geom_col() 
```

```{r most-popular-energy-rap}
spotify_songs %>%
  filter(playlist_genre == "rap",track_popularity > 68) %>%
  summarise(average_energy = mean(energy))
```

Seems like most popular are between 0.50 - 0.75. Not sure how useful it is since we don't have good definition of "energy"


##Check popularity of a song with valence

```{r popularity-vs-valence-rap}
spotify_songs %>%
  filter(playlist_genre == "rap") %>%
  ggplot(aes(y = track_popularity,  x = valence)) +
  geom_col()
```

```{r popularity-vs-valence-by-popularity-lvl-rap}
spotify_songs %>%
  filter(playlist_genre == "rap") %>%
  ggplot(aes(y = track_popularity,  x = valence)) +
  geom_col() +
  facet_wrap(.~ popularity_level)
```

```{r most-popular-valence-rap}
spotify_songs %>%
  filter(playlist_genre == "rap",track_popularity > 68) %>%
  summarise(average_valence = mean(valence))
```

##Check popularity of a song with liveliness INTERESTING

```{r popularity-vs-livileness-rap}
spotify_songs %>%
  filter(playlist_genre == "rap") %>%
  ggplot(aes(y = track_popularity,  x = liveness)) +
  geom_col()
```

```{r most-popular-liveliness-rap}
spotify_songs %>%
  filter(playlist_genre == "rap",track_popularity > 68) %>%
  summarise(average_liveness = mean(liveness))
```
##Check popularity of a song with acousticness

```{r popularity-vs-acousticness-rap}
spotify_songs %>%
  filter(playlist_genre == "rap") %>% 
  group_by(popularity_level) %>%
  summarise(average_acousricness = mean(acousticness))
```

```{r most-popular-accousticness-rap}
spotify_songs %>%
  filter(playlist_genre == "rap",track_popularity > 68) %>%
  summarise(average_accousticness = mean(acousticness))
```


##LATIN

```{r summarry-latin}
spotify_songs %>%
filter(playlist_genre == "latin") %>%
summarise(min_pop = min(track_popularity), iqr_pop = IQR(track_popularity), med_pop = median(track_popularity), max_pop = max(track_popularity), quantile(track_popularity))
```

##Check popularity of a song with key and mode

```{r popularity-vs-key-for-major-latin}
spotify_songs %>%
  filter(playlist_genre == "latin") %>%
  filter(mode == 1) %>%
  group_by(key) %>%
  ggplot(aes(y = track_popularity, x = key)) +
  geom_col() +
  facet_wrap(.~ popularity_level)
```

```{r popularity-vs-key-for-minor-latin}
spotify_songs %>%
  filter(playlist_genre == "latin") %>%
  filter(mode == 0) %>%
  group_by(key) %>%
  ggplot(aes(y = track_popularity, x = key)) +
  geom_col() +
  facet_wrap(.~ popularity_level)
```

c = 0, c# = 1, d = 2, d# = 3,E = 4, F = 5, F# = 6, G = 7, G# = 8, A = 9, A# = 10, B = 11, NA = -1 

##Check popularity of a song with instrumentalness

```{r popularity-vs-instrumentalness-latin}
spotify_songs %>%
  filter(playlist_genre == "latin") %>%
  ggplot(aes(y = track_popularity, x = instrumentalness)) +
  geom_hex() +
  facet_wrap(.~ popularity_level)
```


##Check popularity of a song with duration

```{r popularity-vs-duration-latin}
spotify_songs %>%
  filter(playlist_genre == "latin") %>%
  mutate(duration_minutes = duration_ms/(1000*60)) %>%
  ggplot(aes(x = duration_minutes, y = track_popularity)) +
  geom_point() +
  geom_smooth() 
  
```

```{r most-popular-duration-latin}
spotify_songs %>%
  filter(playlist_genre == "latin",track_popularity > 68) %>%
  summarise(average_duration_latin = mean(duration_ms)/(1000*60))
  
```
 We can see that between 2.5-3.3 there are the most popular songs, which makes sense since very long songs tend not to be so popular.
 
 
##Check popularity of a song with tempo
```{r popularity-vs-tempo-latin}
spotify_songs %>%
  filter(playlist_genre == "latin") %>%
  ggplot(aes(x = tempo, y = track_popularity)) +
  geom_point() +
  geom_smooth() 
```

```{r most-popular-tempo-latin}
spotify_songs %>%
  filter(playlist_genre == "latin",track_popularity > 68) %>%
  summarise(average_tempo = mean(tempo))
  
```

##Check popularity of a song with speechiness

```{r popularity-vs-speechiness-latin}
spotify_songs %>%
  filter(playlist_genre == "latin") %>%
  ggplot(aes(y = track_popularity, x = speechiness)) +
  geom_col() 
```

```{r popularity-vs-speechiness-by-popularity-lvl-latin}
spotify_songs %>%
  filter(playlist_genre == "latin") %>%
  ggplot(aes(y = track_popularity, x = speechiness)) +
  geom_col() +
  facet_wrap(.~ popularity_level)
```


Most popular songs are below 0.25. "Speechiness detects the presence of spoken words in a track. The more exclusively speech-like the recording (e.g. talk show, audio book, poetry), the closer to 1.0 the attribute value. Values above 0.66 describe tracks that are probably made entirely of spoken words. Values between 0.33 and 0.66 describe tracks that may contain both music and speech, either in sections or layered, including such cases as rap music. Values below 0.33 most likely represent music and other non-speech-like tracks."

##Check popularity of a song with mode

```{r popularity-vs-mode-latin}
spotify_songs %>%
  filter(playlist_genre == "latin") %>%
  ggplot(aes(y = track_popularity,  x = mode)) +
  geom_col()
```

```{r popularity-vs-mode-by-popularity-lvl-latin}
spotify_songs %>%
  filter(playlist_genre == "latin") %>%
  ggplot(aes(y = track_popularity,  x = mode)) +
  geom_col() +
  facet_wrap(.~ popularity_level)
```

##Check popularity of a song with danceability

```{r popularity-vs-danceability-latin}
spotify_songs %>%
  filter(playlist_genre == "latin") %>%
  ggplot(aes(y = track_popularity,  x = danceability)) +
  geom_col() 
```
 
```{r popularity-vs-danceability-by-popularity-lvl-latin}
spotify_songs %>%
  filter(playlist_genre == "latin") %>%
  ggplot(aes(y = track_popularity,  x = danceability)) +
  geom_col() +
  facet_wrap(.~ popularity_level)
```

```{r most-popular-danceability-latin}
spotify_songs %>%
  filter(playlist_genre == "latin",track_popularity > 68) %>%
  summarise(average_danceability_latin = mean(danceability))
```


##Check popularity of a song with release date

```{r popularity-vs-releasedate-latin}
spotify_songs %>%
  filter(playlist_genre == "latin") %>%
  mutate(release_date = as_date(track_album_release_date)) %>%
  ggplot(aes(y = track_popularity,  x = release_date)) +
  geom_col()
```

##Check popularity of a song with energy

```{r popularity-vs-energy-latin}
spotify_songs %>%
  filter(playlist_genre == "latin") %>%
  ggplot(aes(y = track_popularity,  x = energy)) +
  geom_col() 
```

```{r most-popular-energy-latin}
spotify_songs %>%
  filter(playlist_genre == "latin",track_popularity > 68) %>%
  summarise(average_energy_latin = mean(energy))
```
Seems like most popular are between 0.50 - 0.75. Not sure how useful it is since we don't have good definition of "energy"


##Check popularity of a song with valence

```{r popularity-vs-valence-latin}
spotify_songs %>%
  filter(playlist_genre == "latin") %>%
  ggplot(aes(y = track_popularity,  x = valence)) +
  geom_col()
```

```{r popularity-vs-valence-by-popularity-lvl-latin}
spotify_songs %>%
  filter(playlist_genre == "latin") %>%
  ggplot(aes(y = track_popularity,  x = valence)) +
  geom_col() +
  facet_wrap(.~ popularity_level)
```

```{r most-popular-valence-latin}
spotify_songs %>%
  filter(playlist_genre == "latin",track_popularity > 68) %>%
  summarise(average_valence_latin = mean(valence))
```
##Check popularity of a song with liveliness INTERESTING

```{r popularity-vs-livileness-latin}
spotify_songs %>%
  filter(playlist_genre == "latin") %>%
  ggplot(aes(y = track_popularity,  x = liveness)) +
  geom_col()
```

```{r most-popular-liveliness-latin}
spotify_songs %>%
  filter(playlist_genre == "latin",track_popularity > 68) %>%
  summarise(average_liveness_latin = mean(liveness))
```


##Check popularity of a song with acousticness

```{r popularity-vs-acousticness-latin}
spotify_songs %>%
  filter(playlist_genre == "latin") %>% 
  group_by(popularity_level) %>%
  summarise(average_acousricness = mean(acousticness))
```

```{r most-popular-accousticness-latin}
spotify_songs %>%
  filter(playlist_genre == "latin",track_popularity > 68) %>%
  summarise(average_accousticness_latin = mean(acousticness))
```

## R&B

```{r summarry-rap-r&b}
spotify_songs %>%
filter(playlist_genre == "r&b") %>%
summarise(min_pop = min(track_popularity), iqr_pop = IQR(track_popularity), med_pop = median(track_popularity), max_pop = max(track_popularity), quantile(track_popularity))
```

##Check popularity of a song with key and mode

```{r popularity-vs-key-for-major-r&b}
spotify_songs %>%
  filter(playlist_genre == "r&b") %>%
  filter(mode == 1) %>%
  group_by(key) %>%
  ggplot(aes(y = track_popularity, x = key)) +
  geom_col() +
  facet_wrap(.~ popularity_level)
```

```{r popularity-vs-key-for-minor-r&b}
spotify_songs %>%
  filter(playlist_genre == "r&b") %>%
  filter(mode == 0) %>%
  group_by(key) %>%
  ggplot(aes(y = track_popularity, x = key)) +
  geom_col() +
  facet_wrap(.~ popularity_level)
```

c = 0, c# = 1, d = 2, d# = 3,E = 4, F = 5, F# = 6, G = 7, G# = 8, A = 9, A# = 10, B = 11, NA = -1 

##Check popularity of a song with instrumentalness

```{r popularity-vs-instrumentalness-r&b}
spotify_songs %>%
  filter(playlist_genre == "r&b") %>%
  ggplot(aes(y = track_popularity, x = instrumentalness)) +
  geom_hex() +
  facet_wrap(.~ popularity_level)
```


##Check popularity of a song with duration

```{r popularity-vs-duration-r&b}
spotify_songs %>%
  filter(playlist_genre == "r&b") %>%
  mutate(duration_minutes = duration_ms/(1000*60)) %>%
  ggplot(aes(x = duration_minutes, y = track_popularity)) +
  geom_point() +
  geom_smooth() 
  
```

```{r most-popular-duration-r&b}
spotify_songs %>%
  filter(playlist_genre == "r&b",track_popularity > 68) %>%
  summarise(average_duration_rnb = mean(duration_ms)/(1000*60))
  
```
 We can see that between 2.5-3.3 there are the most popular songs, which makes sense since very long songs tend not to be so popular.
 
 
##Check popularity of a song with tempo
```{r popularity-vs-tempo-r&b}
spotify_songs %>%
  filter(playlist_genre == "r&b") %>%
  ggplot(aes(x = tempo, y = track_popularity)) +
  geom_point() +
  geom_smooth() 
```

```{r most-popular-tempo-r&b}
spotify_songs %>%
  filter(playlist_genre == "r&b",track_popularity > 68) %>%
  summarise(average_tempo = mean(tempo))
  
```


##Check popularity of a song with speechiness

```{r popularity-vs-speechiness-r&b}
spotify_songs %>%
  filter(playlist_genre == "r&b") %>%
  ggplot(aes(y = track_popularity, x = speechiness)) +
  geom_col() 
```

```{r popularity-vs-speechiness-by-popularity-lvl-r&b}
spotify_songs %>%
  filter(playlist_genre == "r&b") %>%
  ggplot(aes(y = track_popularity, x = speechiness)) +
  geom_col() +
  facet_wrap(.~ popularity_level)
```

```{r most-popular-speechnes-lvl-r&b}
spotify_songs %>%
  filter(playlist_genre == "r&b",track_popularity > 68) %>%
  summarise(average_speechness = mean(speechiness))
```


Most popular songs are below 0.25. "Speechiness detects the presence of spoken words in a track. The more exclusively speech-like the recording (e.g. talk show, audio book, poetry), the closer to 1.0 the attribute value. Values above 0.66 describe tracks that are probably made entirely of spoken words. Values between 0.33 and 0.66 describe tracks that may contain both music and speech, either in sections or layered, including such cases as rap music. Values below 0.33 most likely represent music and other non-speech-like tracks."

##Check popularity of a song with mode

```{r popularity-vs-mode-r&b}
spotify_songs %>%
  filter(playlist_genre == "r&b") %>%
  ggplot(aes(y = track_popularity,  x = mode)) +
  geom_col()
```

```{r popularity-vs-mode-by-popularity-lvl-r&b}
spotify_songs %>%
  filter(playlist_genre == "r&b") %>%
  ggplot(aes(y = track_popularity,  x = mode)) +
  geom_col() +
  facet_wrap(.~ popularity_level)
```

##Check popularity of a song with danceability

```{r popularity-vs-danceability-r&b}
spotify_songs %>%
  filter(playlist_genre == "r&b") %>%
  ggplot(aes(y = track_popularity,  x = danceability)) +
  geom_col() 
```
 
```{r popularity-vs-danceability-by-popularity-lvl-r&b}
spotify_songs %>%
  filter(playlist_genre == "r&b") %>%
  ggplot(aes(y = track_popularity,  x = danceability)) +
  geom_col() +
  facet_wrap(.~ popularity_level)
```

```{r most-popular-danceability-r&b}
spotify_songs %>%
  filter(playlist_genre == "r&b",track_popularity > 68) %>%
  summarise(average_danceability = mean(danceability))
```


##Check popularity of a song with release date

```{r popularity-vs-releasedate-r&b}
spotify_songs %>%
  filter(playlist_genre == "r&b") %>%
  mutate(release_date = as_date(track_album_release_date)) %>%
  ggplot(aes(y = track_popularity,  x = release_date)) +
  geom_col()
```

##Check popularity of a song with energy

```{r popularity-vs-energy-r&b}
spotify_songs %>%
  filter(playlist_genre == "r&b") %>%
  ggplot(aes(y = track_popularity,  x = energy)) +
  geom_col() 
```

```{r most-popular-energy-r&b}
spotify_songs %>%
  filter(playlist_genre == "r&b",track_popularity > 68) %>%
  summarise(average_energy = mean(energy))
```

Seems like most popular are between 0.50 - 0.75. Not sure how useful it is since we don't have good definition of "energy"


##Check popularity of a song with valence

```{r popularity-vs-valence-r&b}
spotify_songs %>%
  filter(playlist_genre == "r&b") %>%
  ggplot(aes(y = track_popularity,  x = valence)) +
  geom_col()
```

```{r popularity-vs-valence-by-popularity-lvl-r&b}
spotify_songs %>%
  filter(playlist_genre == "r&b") %>%
  ggplot(aes(y = track_popularity,  x = valence)) +
  geom_col() +
  facet_wrap(.~ popularity_level)
```

```{r most-popular-valence-r&b}
spotify_songs %>%
  filter(playlist_genre == "r&b",track_popularity > 68) %>%
  summarise(average_valence = mean(valence))
```

##Check popularity of a song with liveliness INTERESTING

```{r popularity-vs-livileness-r&b}
spotify_songs %>%
  filter(playlist_genre == "r&b") %>%
  ggplot(aes(y = track_popularity,  x = liveness)) +
  geom_col()
```

```{r most-popular-liveliness-r&b}
spotify_songs %>%
  filter(playlist_genre == "r&b",track_popularity > 68) %>%
  summarise(average_liveness = mean(liveness))
```
##Check popularity of a song with acousticness

```{r popularity-vs-acousticness-r&b}
spotify_songs %>%
  filter(playlist_genre == "r&b") %>% 
  group_by(popularity_level) %>%
  summarise(average_acousricness = mean(acousticness))
```

```{r most-popular-accousticness-r&b}
spotify_songs %>%
  filter(playlist_genre == "r&b",track_popularity > 68) %>%
  summarise(average_accousticness = mean(acousticness))
```











