---
title: "Exploring the Factors of Song Popularity"
subtitle: "Using TidyTuesdays Spotify dataset"
author: "Louhavidandros <br> Alexandros Zoupos, David Aiton, Harry Rushworth and Louis Parker"
institute: "University of Edinburgh"
date: "`r Sys.Date()`"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      ratio: "16:9"
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
editor_options: 
  chunk_output_type: console
---


```{r load-packages, include = FALSE}
# Add any additional packages you need to this chunk
library(tidyverse)
library(tidymodels)
library(palmerpenguins)
library(knitr)
library(here)
library(patchwork)
library(ggridges)
```

```{r setup, include = FALSE}
# For better figure resolution
knitr::opts_chunk$set(fig.retina = 3, dpi = 300, fig.width = 6, fig.asp = 0.618, out.width = "80%")
```

```{r load-data, include = FALSE}
spotify_songs <- readr::read_csv(here('data/Spotify.csv'))
```

```{r clean-data, include = FALSE}
clean_songs <- spotify_songs %>%
  filter(duplicated(track_name) == FALSE)

clean_full <- spotify_songs %>%
  group_by(playlist_genre) %>%
  filter(duplicated(track_name) == FALSE)
```

```{r popularity-level-clean-data, include = FALSE}
clean_edm <- clean_full %>%
  filter(playlist_genre == "edm") %>%
  mutate(popularity_level = case_when(
    track_popularity <= 17 ~ "Very Low",
    track_popularity > 17 & track_popularity <= 35 ~ "Low",
    track_popularity > 35 & track_popularity < 49 ~ "Fairly High",
    track_popularity >= 49 & track_popularity <= 100 ~ "High"
  ))

clean_latin <- clean_full %>%
  filter(playlist_genre == "latin") %>%
  mutate(popularity_level = case_when(
    track_popularity <= 31 ~ "Very Low",
    track_popularity > 31 & track_popularity <= 48 ~ "Low",
    track_popularity > 48 & track_popularity < 63 ~ "Fairly High",
    track_popularity >= 63 & track_popularity <= 100 ~ "High"
  ))

clean_rnb <- clean_full %>%
  filter(playlist_genre == "r&b") %>%
  mutate(popularity_level = case_when(
    track_popularity <= 21 ~ "Very Low",
    track_popularity > 21 & track_popularity <= 43 ~ "Low",
    track_popularity > 43 & track_popularity < 61 ~ "Fairly High",
    track_popularity >= 61 & track_popularity <= 100 ~ "High"
  ))

clean_rap <- clean_full %>%
  filter(playlist_genre == "rap") %>%
  mutate(popularity_level = case_when(
    track_popularity <= 30 ~ "Very Low",
    track_popularity > 30 & track_popularity <= 47 ~ "Low",
    track_popularity > 47 & track_popularity < 59 ~ "Fairly High",
    track_popularity >= 59 & track_popularity <= 100 ~ "High"
  ))

clean_pop <- clean_full %>%
  filter(playlist_genre == "pop") %>%
  mutate(popularity_level = case_when(
    track_popularity <= 30 ~ "Very Low",
    track_popularity > 30 & track_popularity <= 50 ~ "Low",
    track_popularity > 50 & track_popularity < 65 ~ "Fairly High",
    track_popularity >= 65 & track_popularity <= 100 ~ "High"
  )) 

clean_rock <- clean_full %>%
  filter(playlist_genre == "rock") %>%
  mutate(popularity_level = case_when(
    track_popularity <= 24 ~ "Very Low",
    track_popularity > 24 & track_popularity <= 45 ~ "Low",
    track_popularity > 45 & track_popularity < 61 ~ "Fairly High",
    track_popularity >= 61 & track_popularity <= 100 ~ "High"
  )) 

songs_ridges <- rbind(clean_rock, clean_rnb, clean_rap, clean_pop, clean_latin, clean_edm)
```

#Summary of data

```{r summary_data, echo = FALSE}
summary_table <- spotify_songs %>%
  select(track_name, track_popularity, playlist_genre, danceability) %>%
  slice_max(danceability, n = 5)

kable(head(summary_table), format = "html")
```

---

```{r messy-plot, echo = FALSE, message=FALSE}
clean_full %>%
ggplot(aes(x = speechiness, y = track_popularity)) +
geom_point() +
geom_smooth() +
labs(x = "Speechiness",
y = "Track popularity",
title = "Track popularity vs speechiness") +
theme_minimal() +
scale_color_viridis_d()
```

- Over 27,000 observations
---


```{r data-split, echo = FALSE}
set.seed(666)
clean_songs_split <- initial_split(clean_full)
spotify_train <- training(clean_songs_split)
spotify_test <- testing(clean_songs_split)
```

```{r linear-mod, echo = FALSE}
songs_linear_mod <- linear_reg() %>%
  set_engine("lm")
```

```{r songs-recipe, echo = FALSE}
songs_rec <- recipe(track_popularity ~ ., data = clean_songs) %>%
  update_role(track_name, new_role = "ID") %>%
  update_role(track_artist, new_role = "Artist") %>%
  step_rm(track_id,track_album_id,track_album_name,track_album_release_date,playlist_name,playlist_id,playlist_subgenre) %>%
  step_dummy(all_nominal(), -c(track_name,track_artist)) %>%
  step_zv(all_predictors())
```

```{r workflow, echo = FALSE}
songs_wflow <- workflow() %>%
  add_model(songs_linear_mod) %>%
  add_recipe(songs_rec)
```

```{r fit-for-training, echo = FALSE}
songs_fit <- songs_wflow %>%
  fit(data = spotify_train)
```
Track popularity predicted by model vs actual track popularity

```{r pred-testing, echo = FALSE, message = FALSE}
songs_test_pred <- predict(songs_fit, new_data = spotify_test) %>%
  bind_cols(spotify_test %>% select(track_popularity, track_name, track_artist))

kable(head(songs_test_pred), format = "html")
```

R squared (rsq) and root mean square error (rmse) of model

```{r rsq-rmse, echo = FALSE}
rsquared <- rsq(songs_test_pred, truth = track_popularity, estimate = .pred)

kable(head(rsquared), format = "html")

rmse <- rmse(songs_test_pred, truth = track_popularity, estimate = .pred)

kable(head(rmse), format = "html")
```

---

# Popularity Level

- Every song's popularity level was categorised as either "High", "Fairly High", "Low", or "Very Low" depending on which quartile it fell into.

.pull-left[
```{r quartile-example, echo = FALSE, message = FALSE}
quartile_table <- clean_full %>%
  group_by(playlist_genre) %>%
  summarise(quartiles = quantile(track_popularity)) 

kable(head(quartile_table), format = "html")
```
]
.pull-right[
```{r pop-level table, echo = FALSE}
genre <- c("edm", "edm", "edm", "edm")
popularity <- c("0-17", "18-35", "36-48", "49-99")
popularity_level <- c("Very Low", "Low", "Fairly High", "High")

pop_level_example <- data.frame(genre, popularity, popularity_level)

kable(head(pop_level_example), format = "html")
```
]

---

# Danceability Density Plot - Pop
- Optimal Range = range of danceability values where the red line is higher than the blue
```{r density-plot-example, echo = FALSE}
clean_pop %>%
    filter(popularity_level %in% c("High", "Very Low")) %>%
  ggplot(aes(x = danceability, color = popularity_level)) +
  geom_density(adjust = 2) +
    labs(x = "Danceability",
       y = "Frequency",
       color = "Popularity Level") +
  theme_minimal()
```

---

# Danceability Ridge Plot
```{r ridge-plot-example, echo = FALSE}
songs_ridges %>%
  filter(popularity_level %in% c("High", "Very Low")) %>%
  ggplot(aes(x = danceability, y = playlist_genre, fill = popularity_level, color = popularity_level)) +
   geom_density_ridges(alpha = 0.2, bandwidth = 0.045) +
  labs(x = "Danceability",
       y = NULL,
       fill = "Popularity Level",
       title = "Distibution of Danceability Across Genres",
       subtitle = "Danceability - How suitable a track is for dancing"
  ) + guides(color = FALSE) +
  scale_fill_manual(values = c(
    "Very Low" = "#008dfd",
    "High" = "#fd0000"
  )) +
  theme_minimal()
```

---

# Valence Ridge Plot
```{r valence-ridge, echo = FALSE}
songs_ridges %>%
  filter(popularity_level %in% c("High", "Very Low")) %>%
  ggplot(aes(x = valence, y = playlist_genre, fill = popularity_level, color = popularity_level)) +
   geom_density_ridges(alpha = 0.2, bandwidth = 0.085) +
  labs(x = "Valence",
       y = NULL,
       fill = "Popularity Level",
       title = "Distibution of Valence Across Genres",
       subtitle = "Valence -The positivity or happiness of a track"
       ) +
  guides(color = FALSE) +
  scale_fill_manual(values = c(
    "Very Low" = "#008dfd",
    "High" = "#fd0000"
  )) +
  theme_minimal()
```
---
# Energy Ridge Plot

```{r energy-ridges, echo = FALSE}
songs_ridges %>%
  filter(popularity_level %in% c("High", "Very Low")) %>%
  ggplot(aes(x = energy, y = playlist_genre, fill = popularity_level, color = popularity_level)) +
   geom_density_ridges(alpha = 0.2, bandwidth = 0.07) +
  labs(x = "Energy",
       y = NULL,
       fill = "Popularity Level",
       title = "Distibution of Energy Across Genres",
       subtitle = "Energy - General intensity and level of activity of a track"
       ) +
  guides(color = FALSE) +
  scale_fill_manual(values = c(
    "Very Low" = "#008dfd",
    "High" = "#fd0000"
  )) +
  theme_minimal()

```
---

# Speechiness Ridge Plot

```{r speechiness-ridge, echo = FALSE}
songs_ridges  %>%
filter(popularity_level %in% c("High", "Very Low")) %>%
  ggplot(aes(x = speechiness, y = playlist_genre, fill = popularity_level, color = popularity_level)) +
   geom_density_ridges(alpha = 0.2, bandwidth = 0.04) +
  labs(x = "Speechiness",
       y = NULL,
       fill = "Popularity Level",
       title = "Distibution of Speechiness Across Genres",
       subtitle = "Speechiness - A measure of the presence of spoken (not sung) words"
       ) +
  guides(color = FALSE) +
  scale_fill_manual(values = c(
    "Very Low" = "#008dfd",
    "High" = "#fd0000"
  )) +
  theme_minimal()


```
---

# Acousticness Ridge Plot

```{r acousticness-ridge, echo = FALSE}
songs_ridges %>%
  filter(popularity_level %in% c("High", "Very Low")) %>%
  ggplot(aes(x = acousticness, y = playlist_genre, fill = popularity_level, color = popularity_level)) +
   geom_density_ridges(alpha = 0.2, bandwidth = 0.05) +
  labs(x = "Acousticness",
       y = NULL,
       fill = "Popularity Level",
       title = "Distibution of Acousticness Across Genres",
       subtitle = "Acousticness - A confidence measure of whether the track is acoustic"
       ) +
  guides(color = FALSE) +
  scale_fill_manual(values = c(
    "Very Low" = "#008dfd",
    "High" = "#fd0000"
  )) +
  theme_minimal()
```
---

# Liveness Ridge Plot

```{r liveness-ridge, echo = FALSE}
songs_ridges %>%
  filter(popularity_level %in% c("High", "Very Low")) %>%
  ggplot(aes(x = liveness, y = playlist_genre, fill = popularity_level, color = popularity_level)) +
   geom_density_ridges(alpha = 0.2, bandwidth = 0.045) +
  labs(x = "Liveness",
       y = NULL,
       fill = "Popularity Level",
       title = "Distibution of Liveness Across Genres",
       subtitle = "Liveness - Likelihood the track is a live recording"
       ) +
  guides(color = FALSE) +
  scale_fill_manual(values = c(
    "Very Low" = "#008dfd",
    "High" = "#fd0000"
  )) +
  theme_minimal()
```
---

# Tempo Ridge Plot

```{r tempo-ridge, echo = FALSE}
songs_ridges %>%
  filter(popularity_level %in% c("High", "Very Low")) %>%
  ggplot(aes(x = tempo, y = playlist_genre, fill = popularity_level, color = popularity_level)) +
   geom_density_ridges(alpha = 0.2, bandwidth = 10) +
  labs(x = "Tempo (Bpm)",
       y = NULL,
       fill = "Popularity Level",
       title = "Distibution of Tempo Across Genres",
       subtitle = "Tempo - The estimated tempo of a track in beats per minute"
       ) +
  guides(color = FALSE) +
  scale_fill_manual(values = c(
    "Very Low" = "#008dfd",
    "High" = "#fd0000"
  )) +
  theme_minimal()
```
---

# Duration Ridge Plot

```{r duration-ridge, echo = FALSE}
songs_ridges %>%
  filter(popularity_level %in% c("High", "Very Low")) %>%
  group_by(popularity_level) %>%
  ggplot(aes(x = duration_ms/1000, y = playlist_genre, fill = popularity_level, color = popularity_level)) +
  geom_density_ridges(alpha = 0.2, bandwidth = 20) +
  labs(x = "Duration (s)",
       y = NULL,
       fill = "Popularity Level",
       title = "Distibution of Song Duration Across Genres",
       subtitle = "Duration - A track's length in seconds"
       ) +
  guides(color = FALSE) +
  scale_fill_manual(values = c(
    "Very Low" = "#008dfd",
    "High" = "#fd0000"
  )) +
  theme_minimal()
```
---

# Instrumentalness Ridge Plot

```{r instrumentalness-ridge, echo = FALSE}
songs_ridges %>%
  filter(popularity_level %in% c("High", "Very Low")) %>%
  ggplot(aes(x = instrumentalness, y = playlist_genre, fill = popularity_level, color = popularity_level)) +
   geom_density_ridges(alpha = 0.2, bandwidth = 0.06) +
  labs(x = "Instrumentalness",
       y = NULL,
       fill = "Popularity Level",
       title = "Distibution of Instrumentalness Across Genres",
       subtitle = "Instrumentalness - A measure of the lack of vocals in a track"
       ) +
  guides(color = FALSE) +
  scale_fill_manual(values = c(
    "Very Low" = "#008dfd",
    "High" = "#fd0000"
  )) +
  theme_minimal()
```
---

# Multiple Variable Analysis

```{r pop-multiple-variable-analysis-1, echo = FALSE}
mvg_pop <- clean_pop %>%
  mutate(dance_energy_val = danceability*energy*valence) %>%
  ggplot(aes(x = popularity_level, y = dance_energy_val, colour = popularity_level)) +
  geom_col() +
  labs(
    title = "Pop",
    y = "Count",
    x = "Popularity Level") +
  theme(legend.position = "none")

mvg_edm <- clean_edm %>%
  mutate(dance_energy_val = danceability*energy*valence) %>%
  ggplot(aes(x = popularity_level, y = dance_energy_val, colour = popularity_level)) +
  geom_col() +
  labs(
    title = "EDM",
    y = "Count",
    x = "Popularity Level") +
  theme(legend.position = "none")

mvg_rap <- clean_rap %>%
  mutate(dance_energy_val = danceability*energy*valence) %>%
  ggplot(aes(x = popularity_level, y = dance_energy_val, colour = popularity_level)) +
  geom_col() +
  labs(
    title = "Rap",
    y = "Count",
    x = "Popularity Level") +
  theme(legend.position = "none")

mvg_rock <- clean_rock %>%
  mutate(dance_energy_val = danceability*energy*valence) %>%
  ggplot(aes(x = popularity_level,y = dance_energy_val, colour = popularity_level)) +
  geom_col() +
  labs(
    title = "Rock",
    y = "Count",
    x = "Popularity Level") +
  theme(legend.position = "none")
  

 (mvg_edm | mvg_rap)/(mvg_pop | mvg_rock)
```
---

# Multiple Variable Analysis part 2

```{r multiple-variable-analysis-2, echo = FALSE}
mvg_latin <- clean_latin  %>%
  mutate(dance_energy_val = danceability*energy*valence) %>%
  ggplot(aes(x = popularity_level, y = dance_energy_val, colour = popularity_level)) +
  geom_col() +
  labs(
    title = "Latin",
    y = "Count",
    x = "Popularity Level") +
  theme(legend.position = "none")

mvg_rnb <- clean_rnb %>%
  mutate(dance_energy_val = danceability*energy*valence)  %>%
  ggplot(aes(x = popularity_level, y = dance_energy_val, colour = popularity_level)) +
  geom_col() +
  labs(
    title = "R&B",
    y = "Count",
    x = "Popularity Level") +
  theme(legend.position = "none")

mvg_rnb+mvg_latin
```

---

# Changes in Trends Over Time

```{r Timeline, echo=FALSE, message=FALSE}
songs_ridges %>%
  filter(popularity_level == "High") %>%
  mutate(track_album_release_year = as.integer(str_sub(track_album_release_date, end = 4))) %>%
  filter(track_album_release_year >= 1970) %>%
  group_by(track_album_release_year) %>%
  summarise(acousticness = mean(acousticness), danceability = mean(danceability), energy = mean(energy), liveness = mean(liveness), 
            speechiness = mean(speechiness), valence = mean(valence)) %>%
  pivot_longer(cols = c(danceability, energy, speechiness, acousticness,
         liveness, valence)) %>%
  ggplot(aes(x = track_album_release_year, y = value, colour = name)) +
  geom_point()+
  geom_smooth() +
  theme_minimal() +
  facet_wrap(~name) +
  labs(x = "Release year",
       y = "Value",
       colour = "Variables",
       title = "Mean Value of Variables over Time",
       subtitle = "1970 - 2020") +
theme(axis.text.x = element_text(angle=45, hjust = 1))
```

---
#The models for the variables
```{r create_summary_years, include = FALSE}
mean_year_songs <- songs_ridges %>%
  filter(popularity_level == "High") %>%
  mutate(track_album_release_year = as.integer(str_sub(track_album_release_date, end = 4)))%>%
  filter(track_album_release_year >= 1970) %>%
  select(track_album_release_year, track_popularity, danceability, energy, 
         speechiness, acousticness, instrumentalness, liveness) %>%
  group_by(track_album_release_year) %>%
  summarise(acousticness = mean(acousticness), danceability = mean(danceability), 
            energy = mean(energy), instrumentalness = mean(instrumentalness), 
            liveness = mean(liveness), speechiness = mean(speechiness), popularity = mean(track_popularity))
```

##Danceablility:
```{r models, echo = FALSE}
model <- linear_reg() %>%
  set_engine("lm") %>%
  fit(danceability ~ track_album_release_year, data = mean_year_songs) %>%
  tidy()
kable(head(model), format = "html")
```
- r.squared is 0.4681, noticably larger than our first value.

---
#The next best song of 2021

```{r, echo = FALSE}
model_table <- tibble(acousticness = 0.0874, danceability = 0.6767, 
                      energy = 0.7331, instrumentalness = 0.0097, liveness = 0.1967, valence = 0.8828)
kable(head(model_table), format = "html")
```
---
# Limitations

- Only 6 genres were examined, each of which had 5,000 songs

--

- `track_popularity` variable does not take into account the popularity that songs had at the time they were released

--


- Some songs may have been categorized in the wrong genre
---
class: center
# Thank you for listening


---

