---
title: "Untitled"
output: github_document
---

```{r}
library(tidyverse)
library(broom)
library(here)
library(dplyr)
```

```{r load-data}
spotify_songs <- read_csv(here("data/Spotify.csv"))
```
```{r}
Spotify<- read_csv(here('data/Spotify.csv'))
```


```{r}
Spotify%>%
  filter(playlist_genre=="rock")%>%
  summarise(average=median(mode))
```


```{r}
Spotify_rock <-Spotify%>%
  filter(playlist_genre=="rock")
  
```

```{r}
Spotify_rock%>%
  group_by(mode)%>%
  summarise(avg_pop=mean(track_popularity))
```


```{r}
Spotify_rock%>%
  filter(mode=="0")%>%
  group_by(key)%>%
  summarise(avg_pop=mean(track_popularity))
```

```{r}
Spotify_rock%>%
  filter(mode=="1")%>%
  group_by(key)%>%
  summarise(avg_pop=mean(track_popularity))
```

```{r}
Spotify_rock_pop%>%
  filter(mode=="0")%>%
  group_by(key)%>%
  summarise(avg_pop=mean(track_popularity))
```
```{r}
Spotify_rock_pop%>%
  filter(mode=="1")%>%
  group_by(key)%>%
  summarise(avg_pop=mean(track_popularity))
```


```{r}
Spotify%>%
  filter(playlist_genre=="rock")%>%
  summarise(avg_pop=mean(track_popularity),  avg_danceability=mean(danceability), avg_energy=mean(energy), common_mode=median(mode), avg_speechiness= mean(speechiness), avg_acoustic= mean(acousticness), avg_instrumental= mean(instrumentalness), avg_liveness= mean(liveness), avg_valence= mean(valence), avg_tempo= mean(tempo), avg_length= mean(duration_ms))
```

```{r}
Spotify%>%
  filter(playlist_genre=="rock")%>%
  filter(track_popularity>=70)%>%
  summarise(avg_pop=mean(track_popularity),  avg_danceability=mean(danceability), avg_energy=mean(energy), common_mode=median(mode), avg_speechiness= mean(speechiness), avg_acoustic= mean(acousticness), avg_instrumental= mean(instrumentalness), avg_liveness= mean(liveness), avg_valence= mean(valence), avg_tempo= mean(tempo), avg_length= mean(duration_ms))
```

```{r}
Spotify_rock %>%
summarise(min_pop = min(track_popularity), iqr_pop = IQR(track_popularity), med_pop = median(track_popularity), max_pop = max(track_popularity), quantile(track_popularity))
```

```{r}
    
Spotify_rock <- Spotify_rock %>%
mutate(popularity_level = case_when(
track_popularity <= 21 ~ "Very Low",
track_popularity > 21 & track_popularity <= 46 ~ "Low",
track_popularity > 46 & track_popularity < 63 ~ "Fairly High",
track_popularity >= 63 & track_popularity <= 100 ~ "High"
))

```

```{r}
Spotify_rock %>%
filter(popularity_level == c("High", "Very Low")) %>%
ggplot(aes(x = valence, color = popularity_level)) +
geom_density(adjust = 2) +
labs(x = "Valence",
y = "Frequency",
color = "Popularity Level") +
theme_minimal()
```

```{r}
Spotify_rock %>%
filter(popularity_level == c("High", "Very Low")) %>%
ggplot(aes(x = danceability, color = popularity_level)) +
geom_density(adjust = 2) +
labs(x = "danceability",
y = "Frequency",
color = "Popularity Level") +
theme_minimal()
```

```{r}
Spotify_rock %>%
filter(popularity_level == c("High", "Very Low")) %>%
ggplot(aes(x = tempo, color = popularity_level)) +
geom_density(adjust = 2) +
labs(x = "Tempo",
y = "Frequency",
color = "Popularity Level") +
theme_minimal()
```

```{r}
Spotify_rock %>%
filter(popularity_level == c("High", "Very Low")) %>%
ggplot(aes(x = acousticness, color = popularity_level)) +
geom_density(adjust = 2) +
labs(x = "Acousticness",
y = "Frequency",
color = "Popularity Level") +
theme_minimal()
```


```{r}
Spotify_rock %>%
filter(popularity_level == c("High", "Very Low")) %>%
ggplot(aes(x = speechiness, color = popularity_level)) +
geom_density(adjust = 2) +
labs(x = "Speechiness",
y = "Frequency",
color = "Popularity Level") +
theme_minimal()
```

```{r}
Spotify_rock %>%
filter(popularity_level == c("High", "Very Low")) %>%
ggplot(aes(x = duration_ms, color = popularity_level)) +
geom_density(adjust = 2) +
labs(x = "Duration",
y = "Frequency",
color = "Popularity Level") +
theme_minimal()
```

```{r}
Spotify_rock %>%
  mutate(danceability_energy= danceability*energy)%>%
  filter(popularity_level == c("High", "Very Low")) %>%
  ggplot(aes(x = danceability_energy, color = popularity_level)) +
  geom_density(adjust = 2) +
  labs(x = "Energy",
  y = "Frequency",
  color = "Popularity Level") +
  theme_minimal()
```

```{r}
Spotify_rock %>%
filter(popularity_level == c("High", "Very Low")) %>%
ggplot(aes(x = key, color = popularity_level)) +
geom_density(adjust = 2) +
labs(x = "Key",
y = "Frequency",
color = "Popularity Level") +
theme_minimal()
```


```{r}
Spotify_rock_pop <- Spotify%>%
  filter(playlist_genre=="rock")%>%
  filter(track_popularity>63)%>%
  view()
```

```{r}
Spotify_rock %>%
  
  count(key, sort = TRUE)%>%
  slice_max(n,n=5) %>%
  arrange(key,n)
  
  
```

c = 0, c# = 1, d = 2, d# = 3,E = 4, F = 5, F# = 6, G = 7, G# = 8, A = 9, A# = 10, B = 11



```{r}
Spotify_rock %>%
  filter(mode=="0")%>%
  group_by(key)%>%
  ggplot(aes(x=key,y=track_popularity, color=key, fill=key))+
  geom_col()
```

```{r}
Spotify_rock %>%
  filter(mode=="1")%>%
  group_by(key)%>%
  ggplot(aes(x=key, y=track_popularity, color=key, fill=key))+
  geom_col()
```

```{r}
Spotify_rock_pop %>%
  filter(mode=="0")%>%
  group_by(key)%>%
  ggplot(aes(x=key,y=track_popularity, color=key, fill=key))+
  geom_col()
```

```{r}
Spotify_rock_pop %>%
  filter(mode=="1")%>%
  group_by(key)%>%
  ggplot(aes(x=key,y=track_popularity, color=key, fill=key))+
  geom_col()
```

```{r}
Spotify_rock_pop%>%
  
  ggplot(aes(x=key))+
  geom_histogram(binwidth = 0.5)
  
```



```{r}
Spotify_rock%>%
  ggplot(aes(x=tempo, y=energy, colour=mode))+
  geom_point()+
  geom_smooth()
```

```{r}
Spotify_rock_pop%>%
  ggplot(aes(x=danceability, y=valence, colour=mode))+
  geom_point()
```

```{r}
Spotify_rock_pop%>%
  ggplot(aes(y=track_popularity,x=valence, colour=mode))+
  geom_point()

```







```{r}
rap_songs <- spotify_songs %>%
  filter(playlist_genre == "rap") %>%
  mutate(popularity_level = case_when(
    track_popularity <= 30 ~ "Very Low",
    track_popularity > 30 & track_popularity <= 47 ~ "Low",
    track_popularity > 47 & track_popularity < 60 ~ "Fairly High",
    track_popularity >= 60 & track_popularity <= 100 ~ "High"
  )) 
  
```

```{r}
rnb_songs <- spotify_songs %>%
  filter(playlist_genre == "r&b") %>%
  mutate(popularity_level = case_when(
    track_popularity <= 20 ~ "Very Low",
    track_popularity > 20 & track_popularity <= 44 ~ "Low",
    track_popularity > 44 & track_popularity < 62 ~ "Fairly High",
    track_popularity >= 62 & track_popularity <= 100 ~ "High"
  ))
```

```{r}
latin_songs <- spotify_songs %>%
  filter(playlist_genre == "latin") %>%
  mutate(popularity_level = case_when(
    track_popularity <= 32 ~ "Very Low",
    track_popularity > 32 & track_popularity <= 50 ~ "Low",
    track_popularity > 50 & track_popularity < 67 ~ "Fairly High",
    track_popularity >= 67 & track_popularity <= 100 ~ "High"
  ))
```

```{r}
edm_songs <- spotify_songs %>%
  filter(playlist_genre == "edm") %>%
  mutate(popularity_level = case_when(
    track_popularity <= 17 ~ "Very Low",
    track_popularity > 17 & track_popularity <= 36 ~ "Low",
    track_popularity > 36 & track_popularity < 51 ~ "Fairly High",
    track_popularity >= 51 & track_popularity <= 100 ~ "High"
  )) 
```

```{r popularity-level}
pop_songs <- spotify_songs %>%
  filter(playlist_genre == "pop") %>%
  mutate(popularity_level = case_when(
    track_popularity <= 31 ~ "Very Low",
    track_popularity > 31 & track_popularity <= 52 ~ "Low",
    track_popularity > 52 & track_popularity < 68 ~ "Fairly High",
    track_popularity >= 68 & track_popularity <= 100 ~ "High"
  ))
```


```{r}
Spotify_rock%>%
  mutate(dance_energy_val=danceability*energy*valence)%>%
  ggplot(aes(x=popularity_level,y=dance_energy_val, colour=popularity_level))+
  geom_col()
```


```{r}
pop_songs%>%
  mutate(dance_energy_val=danceability*energy*valence)%>%
  ggplot(aes(x=popularity_level,y=dance_energy_val, colour=popularity_level))+
  geom_col()
```


```{r}
rap_songs%>%
  mutate(dance_energy_val=danceability*energy*valence)%>%
  ggplot(aes(x=popularity_level,y=dance_energy_val, colour=popularity_level))+
  geom_col()
```

```{r}
edm_songs%>%
  mutate(dance_energy_val=danceability*energy*valence)%>%
  ggplot(aes(x=popularity_level,y=dance_energy_val, colour=popularity_level))+
  geom_col()
```

```{r}
latin_songs%>%
  mutate(dance_energy_val=danceability*energy*valence)%>%
  ggplot(aes(x=popularity_level,y=dance_energy_val, colour=popularity_level))+
  geom_col()
```

```{r}
rnb_songs%>%
  mutate(dance_energy_val=danceability*energy*valence)%>%
  ggplot(aes(x=popularity_level,y=dance_energy_val, colour=popularity_level))+
  geom_col()
```
The combination of danceability, valence and energy into one variable shows some interesting results when genres are split into popularity levels. In this case the 3 variables were multiplied together to return a variable that measures how happy, danceable and energetic a song is. As shown in graphs, the songs in the "High" bracket for popularity also had the highest value for this new variable in 4 out of the 6 genres. The 2 genres in which "High" wasn't were R&B and Latin; for Latin however it was a very close second and while the "High" group came last in R&B, this can be explained by the nature of R&B as a genre. In summary while this doesnt explicitly show that the more danceable, energetic and happy the song is the more likely it is to be popular, it does show that popular songs in general, throughout most genres, are more danceable, energetic and happy.
```{r}
Spotify_rock%>%
  mutate(danceability_energy= danceability*energy)%>%
  ggplot(aes(x=danceability_energy, y=valence,colour=popularity_level))+
  
  geom_smooth()
```

```{r}
pop_songs%>%
  mutate(danceability_energy= danceability*energy)%>%
  ggplot(aes(x=danceability_energy, y=valence,colour=popularity_level))+
  
  geom_smooth()
```


```{r}
rnb_songs%>%
  mutate(danceability_energy= danceability*energy)%>%
  ggplot(aes(x=danceability_energy, y=valence,colour=popularity_level))+
  
  geom_smooth()
```

```{r}
latin_songs%>%
  mutate(danceability_energy= danceability*energy)%>%
  ggplot(aes(x=danceability_energy, y=valence,colour=popularity_level))+
  
  geom_smooth()
```

```{r}
edm_songs%>%
  mutate(danceability_energy= danceability*energy)%>%
  ggplot(aes(x=danceability_energy, y=valence,colour=popularity_level))+
  geom_point()+
  geom_smooth()
```

```{r}
rap_songs%>%
  mutate(danceability_energy= danceability*energy)%>%
  ggplot(aes(x=danceability_energy, y=valence,colour=popularity_level))+
  
  geom_smooth()
```

```{r}
spotify_songs%>%
  count(track_artist, sort = TRUE)%>%
  slice_max(n,n=50)
```



```{r}
spotify_songs_count <- spotify_songs%>%
add_count(track_artist)
  
  
```
```{r}
spotify_songs_count<- spotify_songs_count%>%
  filter(n>=46)
```

```{r}
spotify_songs%>%
  group_by(track_artist)%>%
  mutate(avg_pop=mean(track_popularity))%>%
  arrange(desc(avg_pop))%>%
  filter(duplicated(track_artist)==FALSE)%>%
  slice_max(n,n=50)
  
```


```{r}
spotify_songs_count%>%
  group_by(track_artist)%>%
  summarise(track_artist,n,avg_pop=mean(track_popularity),avg_danceability=mean(danceability), avg_energy=mean(energy), common_mode=median(mode), avg_speechiness= mean(speechiness), avg_acoustic= mean(acousticness), avg_instrumental= mean(instrumentalness), avg_liveness= mean(liveness), avg_valence= mean(valence), avg_tempo= mean(tempo), avg_length= mean(duration_ms))%>%
  filter(duplicated(track_artist)==FALSE)
```

```{r}
spotify_songs_count%>%
  summarise(avg_pop=mean(track_popularity),avg_danceability=mean(danceability), avg_energy=mean(energy), common_mode=median(mode), avg_speechiness= mean(speechiness), avg_acoustic= mean(acousticness), avg_instrumental= mean(instrumentalness), avg_liveness= mean(liveness), avg_valence= mean(valence), avg_tempo= mean(tempo), avg_length= mean(duration_ms))
```

```{r}
Spotify%>%
  summarise(avg_pop=mean(track_popularity),  avg_danceability=mean(danceability), avg_energy=mean(energy), common_mode=median(mode), avg_speechiness= mean(speechiness), avg_acoustic= mean(acousticness), avg_instrumental= mean(instrumentalness), avg_liveness= mean(liveness), avg_valence= mean(valence), avg_tempo= mean(tempo), avg_length= mean(duration_ms))
```

```{r}
spotify_songs_count%>%
  group_by(track_artist)%>%
  mutate(avg_pop=mean(track_popularity))%>%
  arrange(desc(avg_pop))%>%
  filter(duplicated(track_artist)==FALSE)
  
```

```{r}
spotify_songs%>%
  group_by(track_artist)%>%
  mutate(avg_pop=mean(track_popularity))%>%
  arrange(desc(avg_pop))%>%
  filter(duplicated(track_artist)==FALSE)
```
None of the 50 most COMMON artists were in the top 50 most "POPULAR" artists


```{r}
spotify_songs_count %>%

ggplot(aes(x = danceability)) +
geom_density(adjust = 2) +
labs(x = "Danceability",
y = "Frequency") +
theme_minimal()
```

```{r}
spotify_songs%>%

ggplot(aes(x = danceability)) +
geom_density(adjust = 2) +
labs(x = "Danceability",
y = "Frequency") +
theme_minimal()
```
```{r}
spotify_songs_count%>%
  group_by(track_artist)%>%
  count(playlist_genre, sort = TRUE)
  #filter(duplicated(playlist_genre)==FALSE)
```
```{r}
spotify_songs_count%>%
  #group_by(track_artist)%>%
  count(playlist_genre, sort = TRUE)
  #filter(duplicated(track_artist)==FALSE)
```
```{r edm multi-variable graph, echo=FALSE }
mvg_edm <- clean_edm%>%
  mutate(dance_energy_val=danceability*energy*valence)%>%
  ggplot(aes(x=popularity_level,y=dance_energy_val, colour=popularity_level))+
  geom_col()
mvg_edm
```

```{r rap multi-variable graph, echo=FALSE}
mvg_rap <- clean_rap%>%
  mutate(dance_energy_val=danceability*energy*valence)%>%
  ggplot(aes(x=popularity_level,y=dance_energy_val, colour=popularity_level))+
  geom_col()
mvg_rap
```

```{r multi-variable graph, echo=FALSE}
(mvg_edm | mvg_rap)/mvg_pop
```