---
title: "Untitled"
output: github_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(tidyverse)
library(dplyr)
library(ggridges)
```

```{r load-data}
spotify_songs <- read_csv("data/Spotify.csv")
```

```{r}
pop_songs <- spotify_songs %>%
  filter(playlist_genre == "pop")
```


### General Summary

```{r general-summary}
pop_songs %>%
  summarise(min_pop = min(track_popularity), iqr_pop = IQR(track_popularity), med_pop = median(track_popularity), max_pop = max(track_popularity), quantile(track_popularity))
```

### Popularity Level

```{r popularity-level}
pop_songs <- pop_songs %>%
  filter(playlist_genre == "pop") %>%
  mutate(popularity_level = case_when(
    track_popularity <= 31 ~ "Very Low",
    track_popularity > 31 & track_popularity <= 52 ~ "Low",
    track_popularity > 52 & track_popularity < 68 ~ "Fairly High",
    track_popularity >= 68 & track_popularity <= 100 ~ "High"
  ))
```

### Median Values

```{r median-values}
pop_songs %>%
  filter(popularity_level == c("High", "Very Low")) %>%
  group_by(popularity_level) %>%
  summarise(med_dance = median(danceability),
            med_energy = median(energy),
            med_tempo = median(tempo),
            med_duration_s = median(duration_ms)/1000,
            med_speech = median(speechiness),
            med_acoustic = median(acousticness)) 
```

### Statistics

```{r dance-energy-stats}
pop_songs %>%
  group_by(popularity_level) %>%
  summarise(min_dance = min(danceability), med_dance = median(danceability), max_dance = max(danceability), min_energy = min(energy), med_energy = median(energy), max_energy = max(energy))
```

```{r duration-acousticness-stats}
pop_songs %>%
  group_by(popularity_level) %>%
  summarise(min_duration_ms = min(duration_ms), med_duration_ms = median(duration_ms), max_duration_ms = max(duration_ms), min_acousticness = min(acousticness), med_acousticness = median(acousticness), max_acousticness = max(acousticness))
```

```{r speechiness-instrumentalness-stats}
pop_songs %>%
  group_by(popularity_level) %>%
  summarise(min_speechiness = min(speechiness), med_speechiness = median(speechiness), max_speechiness = max(speechiness), min_instrumentalness = min(instrumentalness), med_instrumentalness = median(instrumentalness), max_instrumentalness = max(instrumentalness))
```

```{r tempo-stats}
pop_songs %>%
  group_by(popularity_level) %>%
  summarise(min_tempo = min(tempo), med_tempo = median(tempo), max_tempo = max(tempo))
```

### Key Popularity

```{r major-key-histogram}
pop_songs %>%
  filter(mode == "1") %>%
  ggplot(aes(x = key)) +
  geom_histogram() +
  facet_wrap(~popularity_level)
  
```

```{r minor-key-histogram}
pop_songs %>%
  filter(mode == "0") %>%
  ggplot(aes(x = key)) +
  geom_histogram() +
  facet_wrap(~popularity_level)
```

The most common key for a popular pop song with a major key is C# major, followed by C major and then by F# major, whilst the most common keys for an unpopular pop song with a major key are C major then D major and then F# major. Eb major is the least frequent in both popular and unpopular songs, though it is used slighlt more in unpopular songs.

The most common key for a popular pop song with a minor key is C# Minor, followed by B minor and F minor, whilst the most common keys for unpopular pop songs with a minor key are B minor, then A minor and then C# minor
```{r}
major_key_popularity <-pop_songs %>%
  filter(mode == "1") %>%
  group_by(popularity_level, key) %>%
  count(key)
```

``` {r}
minor_key_popularity <- pop_songs %>%
  filter(mode == "0") %>%
  group_by(popularity_level, key) %>%
  count(key)
```

### Most Pop vs Least Pop
```{r}
pop_songs %>%
  filter(popularity_level == c("High", "Very Low")) %>%
  ggplot(aes(x = acousticness, color = popularity_level)) +
  geom_density(adjust = 2)
```

BETWEEN X AND Y MORE LIKELY TO BE V POP THAN V UNPOP

```{r}
pop_songs %>%
  filter(popularity_level == c("High", "Very Low")) %>%
  ggplot(aes(x = danceability, color = popularity_level)) +
  geom_density(adjust = 2)
```

DANCEABILITY IS GOOD

```{r}
pop_songs %>%
  filter(popularity_level == c("High", "Very Low")) %>%
  ggplot(aes(x = energy, color = popularity_level)) +
  geom_density(adjust = 2)
```

ENERGY DECENT

```{r}
pop_songs %>%
  filter(popularity_level == c("High", "Very Low")) %>%
  ggplot(aes(x = loudness, color = popularity_level)) +
  geom_density(adjust = 2)
```

LOUDNESS ALRIGHT - ABOVE X MORE LIKELY TO BE POP

```{r}
pop_songs %>%
  filter(popularity_level == c("High", "Very Low")) %>%
  ggplot(aes(x = speechiness, color = popularity_level)) +
  geom_density(adjust = 2)
```

MEH

```{r}
pop_songs %>%
  filter(popularity_level == c("High", "Very Low")) %>%
  ggplot(aes(x = instrumentalness, color = popularity_level)) +
  geom_density(adjust = 2)
```

WACK

```{r}
pop_songs %>%
  filter(playlist_genre == "pop",
         popularity_level == c("High", "Very Low")) %>%
  ggplot(aes(x = liveness, color = popularity_level)) +
  geom_density(adjust = 2)
```


```{r}
pop_songs %>%
  filter(popularity_level == c("High", "Very Low")) %>%
  ggplot(aes(x = valence, color = popularity_level)) +
  geom_density(adjust = 2)
```

VALENCE USEFUL

```{r}
pop_songs %>%
  filter(popularity_level == c("High", "Very Low")) %>%
  ggplot(aes(x = tempo, color = popularity_level)) +
  geom_density(adjust = 2)
```


```{r}
pop_songs %>%
  filter(popularity_level == c("High", "Very Low")) %>%
  ggplot(aes(x = duration_ms, color = popularity_level)) +
  geom_density(adjust = 2)
```

DURATION ALRIGHT


### New Analysis
```{r}
pop_songs %>%
  group_by(playlist_subgenre) %>%
  ggplot(aes(x = track_popularity, color = playlist_subgenre)) +
  geom_density(adjust = 2)
  
```


```{r}
pop_songs %>%
  group_by(popularity_level) %>%
  count(playlist_subgenre) %>%
  arrange(desc(n))
```



1. Post Teen
2. Dance Pop
3. Electropop
4. Indie

```{r}
pop_songs %>%
  ggplot(aes(x = tempo, y = track_popularity))
```


### Beginning of Nicely Presented Analysis

The quartile values for `track_popularity` were found and these values were used as cut off points to define `popularity_level`, with a song being classified as 'very low', 'low', 'fairly high' and 'high' depending on whether it was higher than any of the quartile values.

The main comparisons are made between songs which have a high popularity level and songs which have a very low popularity level.

```{r immediate-differences}
pop_songs %>%
  filter(popularity_level == c("High", "Very Low")) %>%
  group_by(popularity_level) %>%
  summarise(med_dance = median(danceability),
            med_energy = median(energy),
            med_tempo = median(tempo),
            med_duration_s = median(duration_ms)/1000,
            med_speech = median(speechiness),
            med_acoustic = median(acousticness),
            med_valence = median(valence)) 
```

To begin with, the median values for a selection of variables was compared and the following conclusions were made:
On average, popular pop songs have a higher danceability, speechiness, acousticness and valence than unpopular pop songs, and on average a lower energy and duration than unpopular pop songs. The average tempo for both popular and unpopular pop songs is the same.

### Graphing Popularity Likelihood
Graphical representations then made it easier to tell whether a song was more likely to be popular or unpopular given a certain variable:

### Danceability

```{r danceability}
pop_songs %>%
  filter(popularity_level == c("High", "Very Low")) %>%
  ggplot(aes(x = danceability, color = popularity_level)) +
  geom_density(adjust = 2) +
    labs(x = "Danceability",
       y = "Frequency",
       color = "Popularity Level") +
  theme_minimal()
```

On average, if a pop song has a danceablility over 0.66 it is more likely to be popular than to be very unpopular. Similarly if its has a danceability less than 0.66, it is more likely to be very unpopular than popular. 

Takeaway: Pop songs which are popular are more likely to be suitable for dancing than pop songs which are very unpopular

### Energy

```{r energy}
pop_songs %>%
  filter(popularity_level == c("High", "Very Low")) %>%
  ggplot(aes(x = energy, color = popularity_level)) +
  geom_density(adjust = 2) +
  labs(x = "Energy",
       y = "Frequency",
       color = "Popularity Level") +
  theme_minimal()
```

There is clearly a desirable range of energy values: a pop song with an energy between 0.40 and 0.83 is more likely to be popular than very unpopular, and an energy outwith these values makes a pop song more likely to be very unpopular.

### Valence

```{r valence}
pop_songs %>%
  filter(popularity_level == c("High", "Very Low")) %>%
  ggplot(aes(x = valence, color = popularity_level)) +
  geom_density(adjust = 2) +
    labs(x = "Valence",
       y = "Frequency",
       color = "Popularity Level") +
  theme_minimal()
```

Another optimal range can be seen with valence values:
A pop song with a valence between 0.38 and 0.82 is more likely to be popular than very unpopular.A valence below 0.38 will give a higher likelihood of the pop song being very unpopular, and a valence higher than 0.82 will change the odds in  very slight favour of the song being very unpopular rather than popular.

### Acousticness

```{r acousticness}
pop_songs %>%
  filter(popularity_level == c("High", "Very Low")) %>%
  ggplot(aes(x = acousticness, color = popularity_level)) +
  geom_density(adjust = 2) +
    labs(x = "Acousticness",
       y = "Frequency",
       color = "Popularity Level") +
  theme_minimal()
```

The distribution of acousticness is very similar, but for the biggest likelihood of a pop song being popular rather than unpopular the acousticness should be between 0.07 and 0.63, as other values result in there being a higher likelihood of the song being very unpopular than being popular

### Loudness

```{r}
pop_songs %>%
  filter(popularity_level == c("High", "Very Low")) %>%
  ggplot(aes(x = loudness, color = popularity_level)) +
  geom_density(adjust = 2) +
    labs(x = "Loudness",
       y = "Frequency",
       color = "Popularity Level") +
  theme_minimal()
```

On average, a pop song with a loudness value higher than -7 is more likely to be popular than very unpopular. On average, popular pop songs are louder than very unpopular pop songs

### Duration

```{r}
pop_songs %>%
  filter(popularity_level == c("High", "Very Low")) %>%
  ggplot(aes(x = duration_ms/1000, color = popularity_level)) +
  geom_density(adjust = 2) +
    labs(x = "Duration (s)",
       y = "Frequency",
       color = "Popularity Level") +
  theme_minimal()
```

On average, a pop song with a duration shorter than four minutes (240s) has a greater chance of being popular than unpopular, and durations above four minutes lead to a higher chance of a pop song being unpopular than poular. However the range of durations of pop songs which have the greatest chance of being popular is three minutes to four minutes (180-240s).

### Sub-genre

```{r}
pop_songs %>%
  group_by(playlist_subgenre) %>%
  ggplot(aes(x = track_popularity, color = playlist_subgenre)) +
  geom_density(adjust = 2) +
    labs(x = "Popularity",
       y = "Frequency",
       color = "Sub-Genre") +
  theme_minimal()
```

A post-teen pop song is most likely to have a high popularity out of all of the sub-genres included in the data. Dance pop is the next most likely to have a high popularity, followed by electro pop and then indie poptimism.

An electro pop song is most likely to have a very low popularity out of all the sub-genres included in the data. Indie poptimism is the next most likely to be very unpopular, followed by post-teen pop and dance pop.

*DO RIDGE PLOT OF DISTRIBUTION OF EVERY VARIABLE FOR EACH GENRE*

```{r}
genre_quratiles <- spotify_songs %>%
  group_by(playlist_genre) %>%
  summarise(quantile(track_popularity))
```

```{r}
edm_songs <- spotify_songs %>%
  filter(playlist_genre == "edm") %>%
  mutate(popularity_level = case_when(
    track_popularity <= 17 ~ "Very Low",
    track_popularity > 17 & track_popularity <= 36 ~ "Low",
    track_popularity > 36 & track_popularity < 51 ~ "Fairly High",
    track_popularity >= 51 & track_popularity <= 100 ~ "High"
  )) 
```

```{r}
latin_songs <- spotify_songs %>%
  filter(playlist_genre == "latin") %>%
  mutate(popularity_level = case_when(
    track_popularity <= 32 ~ "Very Low",
    track_popularity > 32 & track_popularity <= 50 ~ "Low",
    track_popularity > 50 & track_popularity < 67 ~ "Fairly High",
    track_popularity >= 67 & track_popularity <= 100 ~ "High"
  ))
```

```{r}
rnb_songs <- spotify_songs %>%
  filter(playlist_genre == "r&b") %>%
  mutate(popularity_level = case_when(
    track_popularity <= 20 ~ "Very Low",
    track_popularity > 20 & track_popularity <= 44 ~ "Low",
    track_popularity > 44 & track_popularity < 62 ~ "Fairly High",
    track_popularity >= 62 & track_popularity <= 100 ~ "High"
  ))
```

```{r}
rap_songs <- spotify_songs %>%
  filter(playlist_genre == "rap") %>%
  mutate(popularity_level = case_when(
    track_popularity <= 30 ~ "Very Low",
    track_popularity > 30 & track_popularity <= 47 ~ "Low",
    track_popularity > 47 & track_popularity < 60 ~ "Fairly High",
    track_popularity >= 60 & track_popularity <= 100 ~ "High"
  )) 
```

```{r}
rock_songs <- spotify_songs %>%
  filter(playlist_genre == "rock") %>%
  mutate(popularity_level = case_when(
    track_popularity <= 21 ~ "Very Low",
    track_popularity > 21 & track_popularity <= 46 ~ "Low",
    track_popularity > 46 & track_popularity < 63 ~ "Fairly High",
    track_popularity >= 63 & track_popularity <= 100 ~ "High"
  )) 
```




```{r}
all_songs <- pop_songs %>%
  rbind(pop_songs, rock_songs, latin_songs, edm_songs, rap_songs, rnb_songs)
```


```{r}
all_songs %>%
  filter(popularity_level == c("High", "Very Low")) %>%
  ggplot(aes(x = valence, y = playlist_genre, fill = popularity_level, color = popularity_level)) +
   geom_density_ridges(alpha = 0.2) +
  labs(x = "Valence",
       y = NULL,
       fill = "Popularity Level"
       ) +
  guides(color = FALSE) +
  scale_fill_manual(values = c(
    "Very Low" = "#008dfd",
    "High" = "#fd0000"
  )) +
  theme_minimal()
```

```{r}
all_songs %>%
  filter(popularity_level == c("High", "Very Low")) %>%
  ggplot(aes(x = danceability, y = playlist_genre, fill = popularity_level, color = popularity_level)) +
   geom_density_ridges(alpha = 0.2) +
  labs(x = "Danceability",
       y = NULL,
       fill = "Popularity Level"
       ) +
  guides(color = FALSE) +
  scale_fill_manual(values = c(
    "Very Low" = "#008dfd",
    "High" = "#fd0000"
  )) +
  theme_minimal()
```

```{r}
all_songs %>%
  filter(popularity_level == c("High", "Very Low")) %>%
  ggplot(aes(x = energy, y = playlist_genre, fill = popularity_level, color = popularity_level)) +
   geom_density_ridges(alpha = 0.2) +
  labs(x = "Energy",
       y = NULL,
       fill = "Popularity Level"
       ) +
  guides(color = FALSE) +
  scale_fill_manual(values = c(
    "Very Low" = "#008dfd",
    "High" = "#fd0000"
  )) +
  theme_minimal()
```

```{r}
all_songs %>%
  filter(popularity_level == c("High", "Very Low")) %>%
  ggplot(aes(x = key, y = playlist_genre, fill = popularity_level, color = popularity_level)) +
   geom_density_ridges(alpha = 0.2) +
  labs(x = "Key",
       y = NULL,
       fill = "Popularity"
       ) +
  guides(color = FALSE) +
  scale_fill_manual(values = c(
    "Very Low" = "#008dfd",
    "High" = "#fd0000"
  )) +
  theme_minimal()
```

```{r}
all_songs %>%
  filter(popularity_level == c("High", "Very Low")) %>%
  ggplot(aes(x = acousticness, y = playlist_genre, fill = popularity_level, color = popularity_level)) +
   geom_density_ridges(alpha = 0.2) +
  labs(x = "Acousticness",
       y = NULL,
       fill = "Popularity"
       ) +
  guides(color = FALSE) +
  scale_fill_manual(values = c(
    "Very Low" = "#008dfd",
    "High" = "#fd0000"
  )) +
  theme_minimal()
```


```{r}
all_songs %>%
  filter(popularity_level == c("High", "Very Low")) %>%
  ggplot(aes(x = speechiness, y = playlist_genre, fill = popularity_level, color = popularity_level)) +
   geom_density_ridges(alpha = 0.2) +
  labs(x = "Speechiness",
       y = NULL,
       fill = "Popularity"
       ) +
  guides(color = FALSE) +
  scale_fill_manual(values = c(
    "Very Low" = "#008dfd",
    "High" = "#fd0000"
  )) +
  theme_minimal()
```

```{r}
all_songs %>%
  filter(popularity_level == c("High", "Very Low")) %>%
  ggplot(aes(x = liveness, y = playlist_genre, fill = popularity_level, color = popularity_level)) +
   geom_density_ridges(alpha = 0.2) +
  labs(x = "Liveness",
       y = NULL,
       fill = "Popularity"
       ) +
  guides(color = FALSE) +
  scale_fill_manual(values = c(
    "Very Low" = "#008dfd",
    "High" = "#fd0000"
  )) +
  theme_minimal()
```


```{r}
all_songs %>%
  filter(popularity_level == c("High", "Very Low")) %>%
  ggplot(aes(x = tempo, y = playlist_genre, fill = popularity_level, color = popularity_level)) +
   geom_density_ridges(alpha = 0.2) +
  labs(x = "Tempo (Bpm)",
       y = NULL,
       fill = "Popularity"
       ) +
  guides(color = FALSE) +
  scale_fill_manual(values = c(
    "Very Low" = "#008dfd",
    "High" = "#fd0000"
  )) +
  theme_minimal()
```

```{r fig.width=20}
all_songs %>%
  filter(popularity_level == c("High", "Very Low")) %>%
  group_by(popularity_level) %>%
  ggplot(aes(x = duration_ms/1000, y = playlist_genre, fill = popularity_level, color = popularity_level)) +
  geom_density_ridges(alpha = 0.2) +
  labs(x = "Duration (s)",
       y = NULL,
       fill = "Popularity Level"
       ) +
  guides(color = FALSE) +
  scale_fill_manual(values = c(
    "Very Low" = "#008dfd",
    "High" = "#fd0000"
  )) +
  theme_minimal()
```

```{r}
all_songs %>%
  filter(popularity_level == c("High", "Very Low")) %>%
  ggplot(aes(x = instrumentalness, y = playlist_genre, fill = popularity_level, color = popularity_level)) +
   geom_density_ridges(alpha = 0.2) +
  labs(x = "Instrumentalness",
       y = NULL,
       fill = "Popularity"
       ) +
  guides(color = FALSE) +
  scale_fill_manual(values = c(
    "Very Low" = "#008dfd",
    "High" = "#fd0000"
  )) +
  theme_minimal()
```


```{r}
edm_quartile_range_all <- all_songs %>%
  mutate(Popularity = case_when(
    track_popularity <= 17 ~ "0-17",
    track_popularity > 17 & track_popularity <= 36 ~ "18-36",
    track_popularity > 36 & track_popularity < 51 ~ "37-50",
    track_popularity >= 51 & track_popularity <= 100 ~ "51-100"
  ))
```

```{r}
genre <- c("edm", "edm", "edm", "edm")
popularity <- c("0-17", "18-36", "37-50", "51-100")
popularity_level <- c("Very Low", "Low", "Fairly High", "High")

pop_level_example <- data.frame(genre, popularity, popularity_level)
```


```{r}
edm_quartile_range_all %>%
  filter(playlist_genre == "edm") %>%
  select(playlist_genre, Popularity, popularity_level) %>%
  group_by(Popularity) %>%
  glimpse()
```


