---
title: "Louis' Practice Document"
author: "Louis"
output: github_document
---

## Setup

```{r install_ggradar, message = FALSE}
devtools::install_github("ricardo-bion/ggradar", 
                          dependencies = TRUE)
```

```{r Setup, message = FALSE}
library(tidyverse)
library(here)
library(ggridges)
library(ggradar)
library(hexbin)
spotify_songs <- readr::read_csv(here('data/Spotify.csv'))
```

```{r define-data, message=FALSE}
edm_songs <- spotify_songs %>%
  filter(playlist_genre == "edm")
feature_names <- names(spotify_songs)[12:23]
```

```{r clean-dataframe}
clean_songs <- spotify_songs %>%
  filter(duplicated(track_name) == FALSE)
clean_edm <- spotify_songs %>%
  filter(playlist_genre == "edm") %>%
  filter(duplicated(track_name) == FALSE)
clean_rock <- spotify_songs %>%
  filter(playlist_genre == "rock") %>%
  filter(duplicated(track_name) == FALSE)
clean_pop <- spotify_songs %>%
  filter(playlist_genre == "pop") %>%
  filter(duplicated(track_name) == FALSE)
clean_rap <- spotify_songs %>%
  filter(playlist_genre == "rap") %>%
  filter(duplicated(track_name) == FALSE)
clean_rnb <- spotify_songs %>%
  filter(playlist_genre == "r&b") %>%
  filter(duplicated(track_name) == FALSE)
clean_latin <- spotify_songs %>%
  filter(playlist_genre == "latin") %>%
  filter(duplicated(track_name) == FALSE)
clean_full <- spotify_songs%>%
  group_by(playlist_genre) %>%
  filter(duplicated(track_name) == FALSE)
```

```{r}
clean_edm <- clean_full %>%
  filter(playlist_genre == "edm") %>%
  mutate(popularity_level = case_when(
    track_popularity <= 17 ~ "Very Low",
    track_popularity > 17 & track_popularity <= 35 ~ "Low",
    track_popularity > 35 & track_popularity < 49 ~ "Fairly High",
    track_popularity >= 49 & track_popularity <= 100 ~ "High"
  ))

clean_latin <- clean_full %>%
  filter(playlist_genre == "latin") %>%
  mutate(popularity_level = case_when(
    track_popularity <= 31 ~ "Very Low",
    track_popularity > 31 & track_popularity <= 48 ~ "Low",
    track_popularity > 48 & track_popularity < 63 ~ "Fairly High",
    track_popularity >= 63 & track_popularity <= 100 ~ "High"
  ))

clean_rnb <- clean_full %>%
  filter(playlist_genre == "r&b") %>%
  mutate(popularity_level = case_when(
    track_popularity <= 21 ~ "Very Low",
    track_popularity > 21 & track_popularity <= 43 ~ "Low",
    track_popularity > 43 & track_popularity < 61 ~ "Fairly High",
    track_popularity >= 61 & track_popularity <= 100 ~ "High"
  ))

clean_rap <- clean_full %>%
  filter(playlist_genre == "rap") %>%
  mutate(popularity_level = case_when(
    track_popularity <= 30 ~ "Very Low",
    track_popularity > 30 & track_popularity <= 47 ~ "Low",
    track_popularity > 47 & track_popularity < 59 ~ "Fairly High",
    track_popularity >= 59 & track_popularity <= 100 ~ "High"
  ))

clean_pop <- clean_full %>%
  filter(playlist_genre == "pop") %>%
  mutate(popularity_level = case_when(
    track_popularity <= 30 ~ "Very Low",
    track_popularity > 30 & track_popularity <= 50 ~ "Low",
    track_popularity > 50 & track_popularity < 65 ~ "Fairly High",
    track_popularity >= 65 & track_popularity <= 100 ~ "High"
  )) 
```


## Including Code

```{r EDM-test}
edm_songs%>%
  select(c('playlist_genre', all_of(feature_names),track_popularity)) %>%
  #filter(track_popularity > 75) %>%
  pivot_longer(cols = feature_names) %>%
  group_by(name)%>%
  ggplot(aes(x = value, colour = name)) +
  geom_density(alpha = 0.5) +
  facet_wrap(~name, scales = "free") +
  labs(title = 'Spotify Audio Feature Density - by Genre',
       x = 'Value', y = 'Density')
```

```{r EDM-top10}
edm_songs%>%
  select(feature_names,track_popularity,track_name) %>%
  slice_max(track_popularity, n = 10) %>%
  pivot_longer(cols = feature_names) %>%
  group_by(name)%>%
  ggplot(aes(x = value, colour = name)) +
  geom_density(alpha = 0.5) +
  facet_wrap(~name, scales = "free")

edm_songs%>%
  select(feature_names,track_popularity,track_name) %>%
  slice_min(track_popularity, n = 10) %>%
  pivot_longer(cols = feature_names) %>%
  group_by(name)%>%
  ggplot(aes(x = value, colour = name)) +
  geom_density(alpha = 0.5) +
  facet_wrap(~name, scales = "free")
```

```{r timeline}
spotify_songs %>%
  mutate(track_album_release_date = as.integer(str_sub(track_album_release_date, end = 4))) %>%
  ggplot(aes(x = track_album_release_date, fill = playlist_genre)) +
  geom_histogram(binwidth = 2)
spotify_songs %>%
  mutate(track_album_release_date = as.integer(str_sub(track_album_release_date, end = 4))) %>%
  group_by(playlist_genre, track_album_release_date) %>%
  summarise(mean_track_popularity =mean(track_popularity)) %>%
  ggplot(aes(x = track_album_release_date, y = mean_track_popularity, colour = playlist_genre)) +
  geom_line() +
  theme_minimal() +
  facet_wrap(~playlist_genre)
```

#quick analysis

The top 10 seem to have much smoother graphs, meaning more variation

## Plan a strategy

I want to create a way of expressing songs based on their characteristics, I believe it's called a radar chart

```{r radar, echo=FALSE}
edm_songs %>%
  slice_max(track_popularity, n = 5) %>%
  group_by(track_name) %>%
  mutate(track_name = str_c(track_name,' ',toString(track_popularity))) %>%
  select(track_name, danceability, energy, key, speechiness, acousticness, instrumentalness,
         liveness, valence) %>%
  mutate(key = key/12) %>%
  ggradar(grid.max = 1)
```

```{r}
mean_year_songs <- clean_pop %>%
  mutate(track_album_release_year = as.integer(str_sub(track_album_release_date, end = 4)))%>%
  select(track_album_release_year, track_popularity, danceability, energy, speechiness, acousticness, instrumentalness,
         liveness) %>%
  group_by(track_album_release_year) %>%
  summarise(acousticness = mean(acousticness), danceability = mean(danceability), 
            energy = mean(energy), instrumentalness = mean(instrumentalness), 
            liveness = mean(liveness), speechiness = mean(speechiness),
            popularity = mean(track_popularity), popularity_diff = 0,acousticness_diff = 0,
            danceability_diff = 0, energy_diff = 0, instrumentalness_diff = 0)
```

```{r}
x <- 1
repeat{
  x <- x + 1
  mean_year_songs$popularity_diff[x] <- (mean_year_songs$popularity[x] - mean_year_songs$popularity[x-1])/100
  mean_year_songs$acousticness_diff[x] <- mean_year_songs$acousticness[x] - mean_year_songs$acousticness[x-1]
  mean_year_songs$danceability_diff[x] <- mean_year_songs$danceability[x] - mean_year_songs$danceability[x-1]
  mean_year_songs$energy_diff[x] <- mean_year_songs$energy[x] - mean_year_songs$energy[x-1]
  mean_year_songs$instrumentalness_diff[x] <- mean_year_songs$instrumentalness[x] - mean_year_songs$instrumentalness[x-1]
  if(x==nrow(mean_year_songs)){break
    }
}
```

```{r}
mean_year_songs %>%
  filter(track_album_release_year > 1965) %>%
  select(track_album_release_year,popularity_diff,acousticness_diff,danceability_diff,energy_diff,instrumentalness_diff) %>%
  pivot_longer(cols = c(acousticness_diff, danceability_diff, energy_diff, instrumentalness_diff)) %>%
  group_by(name) %>%
  ggplot(aes(x = name, fill = value * popularity_diff > 0)) +
  geom_bar(position = "dodge")
```

This tracks the correlation of year to year changes of each variable to the year to year popularity changes and tallies whether an increase in a variable happened at the same time as an increase in popularity.
